---
title: "Confidence Intervals -- Real Life Sampling Distributions"
date: "November 1, 2021"
author: "Dr. Allison Theobold"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "slide-style.css"]
    nature:
      highlightStyle: solarized-light
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# R options
options(
  htmltools.dir.version = FALSE,
  tibble.width = 65,
  width = 65
  )

# figure height, width, dpi
knitr::opts_chunk$set(fig.width = 8, 
                      fig.asp = 0.618,
                      out.width = "60%",
                      dpi = 300, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center",
                      echo = FALSE)

# fontawesome
htmltools::tagList(rmarkdown::html_dependency_font_awesome())

# magick
dev.off <- function(){
  invisible(grDevices::dev.off())
}

# xaringanExtra
library(xaringanExtra)
xaringanExtra::use_panelset()

library(emo)
library(tidyverse)
library(openintro)
library(ggridges)
library(flair)
library(broom)
library(gridExtra)
library(kableExtra)
library(png)
library(moderndive)
library(infer)
library(scales)

options(show.signif.stars = FALSE)
```

```{r set-theme, include = FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color      = "#b76352", # mango
  secondary_color    = "#34605f", # bayberry
  header_font_google = google_font("Raleway"),
  text_font_google   = google_font("Raleway", "300", "300i"),
  code_font_google   = google_font("Source Code Pro"),
  header_color = "#793540", #rhubarb
  white_color = "#F5F5F5", # lightest color
  black_color = "#36454F", # darkest color
  text_font_size = "30px", 
  link_color = "#696969" #grey
)
```

class: center, middle, inverse

.hand[.huge[What if I only have one sample?]]

---

class: center, middle 

.larger[Approximate the variability you'd expect to see in other samples!] 

<!-- make a connection between sampling distributions and the approximations we use -->

--

.larger[**Bootstrapping!**]

---

class: middle 

.larger[A Bootstrap Resample]

--

Assumes the original sample is "representative" of observations in the
population. 

--

Uses the original sample to generate new samples that might have occurred with 
additional sampling. 

--

.dark-blue[We can use the statistics from these bootstrap samples to approximate
the true sampling distribution!]

---

class: center, middle, inverse

.hand[.huge[Why???]]

---

class: middle

.larger[Estimating a population parameter]

--

.grey[We are interested in knowing how a statistic varies from sample to 
sample.]

--

.bayberry[Knowing a statistic's behavior helps us make better / more informed
decisions!]

--

.rhubarb[This helps us estimate what values are more or less likely for the 
population parameter to have.]

---

class: inverse, center

.huge[Confidence Intervals]

--

.large[Capture a range of plausible values for the population parameter.]

--

.large[Are more likely to capture the population parameter than a *point
estimate*.]

---

class: middle

.larger[Using bootstrap resamples to generate a confidence interval]

--

From your original sample, resample with replacement the same number of times 
as your original sample. 

--

<center>
This is your .pink[bootstrap resample]. 
</center>

--

Repeat this process many, many times. 

--

Calculate a numerical summary (e.g., mean, median) for each bootstrap 
resample. 

--

<center>
These are your .pink[bootstrap statistics]. 
</center>


---

class: inverse, center

.larger[Bootstrap Distribution]

--

.large[*definition*: a distribution of the .pink[bootstrap statistics] from every .pink[bootstrap resample]].

--

.large[Displays the variability in the statistic that could have happened with 
repeated sampling.] 

--

.large[.mango[Approximates the true sampling distribution!]]
