---
title: "Concept Questions for SLR"
format: html
editor: visual
---

`r emo::ji("scroll")` Group Concept Questions

4.  Below is a regression table for the linear regression model of evaluation score as a function of the age of professor. Use the regression table to write out the estimated regression line for this relationship.

```{r, echo = FALSE}
age_lm <- lm(score ~ age, data = evals) 
get_regression_table(age_lm)
```

5.  Below is information on one professor from the `evals` dataset. Compute the residual for this professor based on the regression line you found in #4.

```{r, echo = FALSE}
evals %>% 
  select(ID, score, age, bty_avg) %>% 
  sample_n(size = 1)
```

6.  What does the term "offset" mean in a linear regression with a categorical explanatory variable?

7.  The `evals` data has a variable named `ethnicity` with levels `"minority"` and `"not minority"`. Using the table below, write down the estimated mean evaluation score for professors of these two different ethnicity groups.

```{r, echo = FALSE}
ethnic_lm <- lm(score ~ ethnicity, data = evals)
get_regression_table(ethnic_lm)
```

8.  What procedure / criteria does `R` use to decide what constitutes the "best" regression line? How would you describe this procedure to someone?
