---
title: "Concept Questions for SLR"
format: html
editor: visual
---


`r emo::ji("scroll")` Group Concept Questions 

1.	What are two types of data modeling? What purpose does each serve? What type
of data modeling have you seen before? 

2.	The plot below suffers from what we call "overplotting". Describe what overplotting is and how it can be remedied. 

```{r, echo = FALSE}
ggplot(evals, aes(x = bty_avg, y = score)) +
  geom_point() +
  labs(x = "Beauty Score", 
       y = "Teaching Score",
       title = "Scatterplot of relationship of teaching and beauty scores")
```

 
3. Below is a plot of the relationship between the age of the professor and
their teaching evaluations. 
Describe the relationship you see.
Does the relationship between age and evaluation scores appear to be stronger or weaker than the relationship with beauty scores? 

```{r, echo = FALSE}
ggplot(evals, aes(x = age, y = score)) +
  geom_jitter() +
  labs(x = "Age of Professor", 
       y = "Teaching Score",
       title = "Scatterplot of relationship of teaching scores and age")
```

4. Below is a regression table for the linear regression model of evaluation
score as a function of the age of professor.
Use the regression table to write out the estimated regression line for this relationship. 

```{r, echo = FALSE}
age_lm <- lm(score ~ age, data = evals) 
get_regression_table(age_lm)
```

5. Below is information on one professor from the `evals` dataset.
Compute the residual for this professor based on the regression line you found
in #4. 

```{r, echo = FALSE}
evals %>% 
  select(ID, score, age, bty_avg) %>% 
  sample_n(size = 1)
```

6. What does the term "offset" mean in a linear regression with a categorical explanatory variable? 

7. The `evals` data has a variable named `ethnicity` with levels `"minority"`
and `"not minority"`. 
Using the table below, write down the estimated mean evaluation score for
professors of these two different ethnicity groups. 

```{r, echo = FALSE}
ethnic_lm <- lm(score ~ ethnicity, data = evals)
get_regression_table(ethnic_lm)
```

8. What procedure / criteria does `R` use to decide what constitutes the "best" regression line?
How would you describe this procedure to someone? 

