---
title: "The Future of Science (Hopefully)"
date: "October 21, 2021"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "slide-style.css"]
    nature:
      highlightStyle: solarized-light
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---


```{r, echo = FALSE, message = FALSE, warning = FALSE}
# R options
options(
  htmltools.dir.version = FALSE,
  tibble.width = 65,
  width = 65
  )

# figure height, width, dpi
knitr::opts_chunk$set(fig.width = 8, 
                      fig.asp = 0.618,
                      out.width = "70%",
                      dpi = 300, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center", 
                      echo = FALSE)

# fontawesome
htmltools::tagList(rmarkdown::html_dependency_font_awesome())

# magick
dev.off <- function(){
  invisible(grDevices::dev.off())
}

# xaringanExtra
library(xaringanExtra)
xaringanExtra::use_panelset()

library(emo)
library(tidyverse)
library(openintro)
library(gridExtra)
library(kableExtra)
library(moderndive)

options(show.signif.stars = FALSE)
```

```{r set-theme, include = FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color      = "#b76352", # mango
  secondary_color    = "#34605f", # bayberry
  header_font_google = google_font("Raleway"),
  text_font_google   = google_font("Raleway", "300", "300i"),
  code_font_google   = google_font("Source Code Pro"),
  header_color = "#793540", #rhubarb
  white_color = "#F5F5F5", # lightest color
  black_color = "#36454F", # darkest color
  text_font_size = "30px", 
  link_color = "#696969" #grey
)
```

class: center, middle, inverse

.huge[.hand[Humans interpret data]]

---

.larger[A familiar context]

```{r, eco}
table <- MA_schools %>% 
  lm(average_sat_math ~ perc_disadvan, data = .) %>% 
  get_regression_table() %>% 
  select(term, estimate, std_error)

table %>% 
  kable() %>% 
  kableExtra::kable_styling()
```

For a 1% increase in students who are economically disadvantaged, we expect the
mean SAT math score to decrease by approximately `r filter(table, term == "perc_disadvan")$estimate`. 

---

.larger[Your project]

```{r}
MA_schools %>% 
  ggplot(mapping = aes(x = perc_disadvan, y = average_sat_math, 
                       color = size)) + 
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm", alpha = 0.5) + 
  labs(x = "Percentage of Students Economically Disadvantaged", 
       y = "Average SAT Math Score", 
       color = "Size of School")
```

---

class: middle
.larger[**Hermeneutics**  hərmə'n(y)o͞odiks]

> *the study of interpretation*

</br>
--

.bitlarger[The philosophy of interpretation]

<!-- "Hermeneutics treats interpretation itself as the subject matter and not as  -->
<!-- auxiliary to the study of something else."  -->

> "Philosophically, hermeneutics concerns the meaning of interpretation--its
> basic nature, scope and validity, and its place within and implications for
> human existance." 

---

.larger[**Statistical Hermeneutics**]

> "The study of the interpretation of results and conclusions based on
> statistical theory, methods, and reasoning."

--

- The complexities in interpretations of statistical models, results, etc. are
often downplayed and taken for granted. 

- It is not often that we are hit with an *explicit* call to discuss our 
*implicit* interpretations of the statements that arise from our use of
statistical methods and results. 

---

class: middle, inverse

.larger[Better for Science]

> "Modern science despite all of the methodological and technological 
> sophistication, has failed to account for the basic epistemic foundation 
> on which it lies." 

--

- Hermeneutics possesses a positive attitude toward the finitude of human 
understanding. 

  * The concern is not to eradicate our finitude, but to recognize the 
  consequences of our limits. 
  
--

> "This finitude is itself an important opening for the persuit of new and 
> different meaning." 
