---
output: pdf_document
---


```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 8, 
                      fig.asp = 0.618,
                      out.width = "70%",
                      dpi = 300, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center", 
                      echo = FALSE)

library(emo)
library(tidyverse)
library(openintro)
library(gridExtra)
library(moderndive)

options(show.signif.stars = FALSE)
```

<!-- Give example of report with same slopes but no justification.  -->

This study seeks to investigate how the relationship between average SAT math
scores and the percentage of students who are economically disadvantaged 
differs based on the size of the school. As seen in the plot below, the slope
for this relationship was found to be similar across the different levels of
schools. Therefore, this study used a parallel slopes (additive) multiple linear
regression model.

```{r}
MA_schools %>% 
  ggplot(mapping = aes(x = perc_disadvan, y = average_sat_math, 
                       color = size)) + 
  geom_point(alpha = 0.5) + 
  geom_parallel_slopes(alpha = 0.5) + 
  labs(x = "Percentage of Students Economically Disadvantaged", 
       y = "Average SAT Math Score", 
       color = "Size of School")
```

\vspace{0.5in}

\begin{center}

\textbf{What do you feel is missing from this model justification?}

\end{center}