---
title: "Multiple Regression with a Numerical & Categorical Variable"
output: ioslides_presentation
fig_height: 4
fig_width: 4
---

<style type="text/css">
slides > slide:not(.nobackground):after {
  content: '';
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                       message = FALSE, 
                      fig.width = 5, 
                      fig.height = 4, 
                      results = "hold")

options(show.signif.stars = FALSE)
library(tidyverse)
library(moderndive)
library(openintro)
library(flair)
library(broom)
library(gridExtra)
library(kableExtra)

evals <- evals %>% 
  rename(sex = gender)
```

## Simple Linear Regression Review

We can model the relationship between one quantitative explanatory variable and 
one response variable with a "simple" linear regression. 

For example, we could model a professor's evaluation score (`score`)
and their average attractiveness score (`bty_avg`) with a simple linear 
regression!

```{r, echo = FALSE, fig.width = 4, fig.height = 3, fig.align = 'center'}
evals %>% 
  ggplot(aes(x = bty_avg, y = score)) + 
  geom_jitter() + 
  geom_smooth(method = "lm") +
  labs(x = "Average Beauty Score", 
       y = "Course Evaluation Score")
  
```

## Including a Categorical Variable

What if we have groups? 

```{r, echo = FALSE, fig.width = 6, fig.align = 'center'}
evals %>% 
  ggplot(aes(x = bty_avg, y = score, color = sex)) + 
  geom_jitter() + 
  geom_smooth(method = "lm") +
  geom_abline(slope = 0.06664, intercept = 3.88034, color = "black", lty = 2) + 
  labs(x = "Average Beauty Score", 
       y = "Course Evaluation Score", 
       color = "Sex of Professor")
  
```