---
title: "Grading Guide - Lab 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1 (a)
How large is the `starwars` dataset? (i.e. How many rows and columns does it have?)

**4**: 

- Must have `glimpse(starwars)` somewhere in their code
- Must state there are 87 rows and 14 columns

**Redo**: Don't mention how many rows or columns there are


# Question 1 (b)
Are there categorical variables in the dataset? If so, what are their names?  

**4**: 

- Lists all variables labeled as `fct` or `chr`
  * name, hair_color, skin_color, eye_color, sex, gender, homeworld, species

**3**: 

- Has all but 1 or 2 of the variables

**Redo**: 

- If their response includes numeric variables
- If their response includes list variables (films, vehicles, starships)


# Question 2
What are the distinct levels / values of the `gender` variable? 

**4**: 

- Provides code that uses `distinct()` function
- States there are **3** levels: masculine, feminine, and NA

**3**:

- Provides code that uses `distinct()` function
- States there are **2** levels: masculine and feminine (misses NA level)

**Redo**: 

- Does not use `distinct()` to find levels


# Question 3
Use the `mutate()` and `if_else()` functions to create a new variable named
`human` based on the criteria outlined above. 

**4**: Code should look similar to:

```{r, eval = FALSE}
starwars <- mutate(starwars, 
                   human = if_else(species == "Human", 
                                   "human", 
                                   "not human")
```

- Can use ` %>%` or pass in data as first argument
- Leave a note if they use an `=` instead of an `<-`

# Question 4 
Using the new `human` variable, create a new dataset named
`human_characters` that only contains observations where the value of `human` 
is equal to "human". 

**4**: Code should look similar to:

```{r, eval = FALSE}
human_characters <- filter(starwars, 
                           human == "human")
```

- Can use ` %>%` or pass in data as first argument
- Leave a note if they use an `=` instead of an `<-`

# Question 5
Take your code from problems 3 and 4 and combine them so that the two steps
are connected with a pipe operator.

**4**: Code should look similar to:

```{r, eval = FALSE}
human_characters <- starwars %>% 
  mutate(human = if_else(species == "Human", 
                         "human", 
                         "not human")) %>% 
  filter(human == "human")
```

**3**: 

- If the linking between the pipes is not correct 
- If they create a new dataset (not `human_characters`)

**Redo**: 

- If their code doesn't run

# Question 6 
Now, using the `human_characters` dataset you made, summarize the number 
of characters from each `homeworld`. 

**4**: 

- Uses the human_characters dataset
- Uses the `count()` function to summarize homeworld

**3**:

- Uses the starwars dataset
- Uses the `count()` function to summarize homeworld

**Redo**: 

- Uses anything other than `count()` to summarize homeworld

# Question 7 
Create a visualization of the summaries you found in question 6. 

**4**: 

- Uses the human_characters dataset
- Uses ggplot to create a barplot
- Displays the homeworld on **y** axis 

**3**:

- Uses ggplot to create a barplot
- Displays the homeworld on **x** axis 

**Redo**: If any of the following are true:

- Uses the starwars dataset
- Uses something other than ggplot to create plot
