---
title: "Lab 9 -- One-Way ANOVA"
format: 
  html:
    self-contained: true
editor: visual
execute: 
  echo: true
  message: false
  warning: false
---


```{r packages}
library(tidyverse)
library(infer)
```

## Today's Data

```{r data-cleaning}

math <- read_csv(here::here("Week 9 - ANOVA",
                            "Lab",
                            "data", 
                            "math_achievement_8th_grade.csv")
                 ) %>% 
  pivot_longer(cols = -country, names_to = "year", values_to = "grade_8_math_score") %>% 
  drop_na(grade_8_math_score)

countries_continents <- read_csv("https://raw.githubusercontent.com/dbouquin/IS_608/master/NanosatDB_munging/Countries-Continents.csv")


math_continents <- math %>% 
  left_join(countries_continents, by = c("country" = "Country")
            ) %>% 
  rename(continent = Continent) %>% 
  mutate(continent = case_when(country == "Czech Republic" ~ "Europe", 
                               country == "Russia" ~ "Asia", 
                               country == "South Korea" ~ "Asia", 
                               country == "United States" ~ "North America", 
                               .default = continent), 
         year_fct = as.factor(year)
         )
         
```

These data come from the Gapminder Foundation, an organization interested in 
increasing the use and understanding of statistics and other information about
social, economic and environmental development at local, national and global
levels.

- Math Achievement -- Average score in the international TIMSS test for children 
in the 8th grade 

We will be analyzing data form 42 countries from across the world from 2007. 


## Data Visualizations

## Statistical Model

### Model Conditions

## One-Way ANOVA Inference

### Testing for a Difference Between Years



### Testing for a Difference Between Continents


### Testing for a Difference Between Continents AND Years