---
title: "Week Five: Diving Deeper into Regression (Multiple Linear Regression)"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
library(gapminder)
library(flair)
library(emo)
library(openintro)
library(moderndive)
library(skimr)
library(gridExtra)
```


Welcome!

In this week's coursework we are going to build off the concepts we learned 
last week and delve deeper into linear regression. We are going to explore 
*multiple linear regression*, a statistical model where we have multiple 
explanatory variables and a single numerical response. We are going to refresh
how to visualize these types of models, practice fitting these types of models 
in `R`, and the learn how to interpret these types of models. 

## Learning Outcomes 

By the end of this coursework you should be able to:

- describe to someone what a multiple linear regression is
- outline how a categorical explanatory variable can be included in a simple 
linear regression
- visualize multiple linear regression models with one numerical and one categorical explanatory variable
- calculate the simple linear regression line for each group of a categorical 
variable
- outline how a second numerical explanatory variable can be included in a 
simple linear regression
- visualize multiple linear regression models with two  numerical explanatory
variables
- interpret the coefficient of each explanatory variable included in the 
regression model
- recite different methods that can be used to decide what multiple linear 
regression model is "best"
- describe the benefits and costs of using "model selection" for deciding on 
a multiple linear regression model


## This Week's TODOs

`r emo::ji("book")` Reading: 45-60 minutes 

`r emo::ji("paper")` Group Discussion Homework: MLR Project  

`r emo::ji("check")` Check-ins: 0 (Group Project Week)

---

# Multiple Regression  

`r emo::ji("book")` [**Required Reading:** *Modern Dive*, Multiple Regression](https://moderndive.com/6-multiple-regression.html)

---

# Group Project

**Details will be announced in class on Monday and posted here afterward.** 
