---
title: "Interpreting an ANOVA"
output: ioslides_presentation
---

<style type="text/css">
slides > slide:not(.nobackground):after {
  content: '';
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE)
library(tidyverse) 
library(palmerpenguins)
library(moderndive)
library(openintro)
library(broom)
library(ggridges)
```

## Difference in Penguin Bill Length 

- We're interested in testing to see if the mean bill depth differs across
species of penguins in the Palmer Archipelago. 

```{r, fig.width=6, fig.height=4, fig.align='center', echo = FALSE}
penguins %>% 
  ggplot(aes(y = species, x = bill_depth_mm, fill = species)) + 
  geom_density_ridges(alpha = 0.8) + 
  theme(legend.position = "none") + 
  labs(x = "Bill Depth (mm)", 
       y = "Species of Penguin")
```

## Model Conditions 

- It seems reasonable to assume that the penguins within each species are
independent. 

- These penguins were not randomly sampled, but it seems reasonable to assume 
that the observations within a group are independent. 
  <!-- * The sample sizes (152, 68, and 124) are certainly less than 10% of the -->
  <!-- population.  -->

- The distributions have some oddities, but not enough to conclude they violate 
the normality condition. 

- The shapes of the distribution appear relatively similar.
  * The Adelie penguins have more variability in their bill depths, but nothing 
  too outlandish. 
  
## Conducting an ANOVA $F$-test

```{r}
aov(bill_depth_mm ~ species, data = penguins) %>% 
  tidy()
```

## Obtaining ANOVA Coefficients

```{r}
penguins_lm <- lm(bill_depth_mm ~ species, data = penguins)

get_regression_table(penguins_lm)
```

## Interpreting ANOVA Coefficients 

```{r, echo = FALSE}
get_regression_table(penguins_lm)
```

</br>

$$ \overline{bill depth}_{Adelie} = 18.3 \text{mm}$$ 

$$ \overline{bill depth}_{Chinstrap} = 18.3 + 0.074 = 18.374 \text{mm}$$ 

$$ \overline{bill depth}_{Gentoo} = 18.3 - 3.36 = 14.94 \text{mm}$$ 




