---
title: "ANOVA Visualizations"
output: ioslides_presentation
---

<style type="text/css">
slides > slide:not(.nobackground):after {
  content: '';
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE)
library(tidyverse) 
library(palmerpenguins)
library(gridExtra)
library(openintro)
library(skimr)
library(infer)
library(okcupiddata)
```

## ANOVA

Analysis of variance (ANOVA) compares the means of three of more groups to 
detect if the means of the groups are different.  

</br>

We want visualizations that allow for us to easily compare:  

- the center (mean) of the groups 
- the spread (variability) of the groups 


## Data Visualizations 

__Side-by-Side Violin Plots__ 

```{r, fig.width=4, fig.height=3, fig.align='center'}
penguins %>% 
  ggplot(aes(x = bill_length_mm, y = island)) + 
  geom_violin(aes(fill = island)) + 
  geom_jitter() +
  theme(legend.position = "none")
```

## Faceted Boxplots 

```{r, fig.width=5, fig.height=4, fig.align='center'}
profiles %>% 
  filter(sex == "m") %>% 
  ggplot(aes(x = height)) +
  geom_histogram() + 
  facet_wrap(~orientation)
```


