---
title: "Code Handout -- Week 2"
author: "Allison Theobold"
output: pdf_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(fig.width = 3, 
                  fig.height = 3, 
                  message = FALSE, 
                  warning = FALSE,
                  eval = FALSE)

```

This document contains all of the functions that we have covered thus far in the
course. It will be updated every week, after we've added new skills. Each 
function is presented alongside an example of how it is used. 

All of the examples below are in the context of the Palmer Penguins, found 
[here (link)](https://allisonhorst.github.io/palmerpenguins/index.html). 

\vspace{0.5cm}

## Packages

- `library()` -- loads packages into your `R` session
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(moderndive)
library(skimr)
library(broom)

library(palmerpenguins) 
```


## Other Data Wrangling Tools


  
\vspace{0.5cm}

- `slice_sample()` -- selects rows from the dataframe, based on the value of 
`n` specified

```{r}
penguins %>% 
  slice_sample(n = 10)
```

\vspace{0.5cm}

