---
title: "Untitled"
format: docx
editor: visual
---

- `c()` -- the "concatenate" function combines inputs to form a vector, the 
values have to be the same data type. 

```{r}
cat_variables <- c("Species", "Island", "Sex")
```


- `group_by()` -- groups the dataframe based on levels of a categorical 
variable (usually used alongside `summarize()` or `count()`)

```{r, eval = FALSE}
penguins %>% 
  group_by(island)
```

- `count()` -- counts the number of observations (rows) in different levels of a 
categorical variable (can be used with `group_by()` to include two categorical 
variables) 

```{r}
penguins %>% 
  count(island)

penguins %>% 
  group_by(island) %>% 
  count(species)
```


- `if_else()` -- function which creates output based on a condition to be 
satisfied, the first argument is the logical test we wish to perform, the second
argument is what we want output if the result of the logical test is `TRUE`,
the third argument is what we want output if the result of the logical test is
`FALSE`
```{r}
penguins %>%
  filter(is.na(flipper_length_mm) != TRUE) %>% 
  mutate(large_flipper = 
           if_else(flipper_length_mm >= mean(flipper_length_mm),
                   "above average",
                   "below average")
         )
```

\vspace{0.5cm}

- `as.factor()` -- converts a variable from numerical or character into a factor, 
necessary if fitting a data model or producing a visualization with a numerical 
variable that you want treated as categorical
  * `as.character()` will also work! 

```{r}
penguins %>% 
  mutate(year_chr = as.character(year), 
         year_fct = as.factor(year))
```

  


# Faceted Histograms

ggplot(data = <NAME OF DATASET>, 
       mapping = aes(x = <NAME OF NUMERICAL VARIABLE>)) + 
  geom_histogram(binwidth = <WIDTH OF BINS>) + 
  facet_wrap(~<NAME OF CATEGORICAL VARIABLE>) +
  labs(x = "<TITLE FOR THE X-AXIS>")

Note: A histogram must have the variable on the x-axis! 

Side-by-Side Boxplots

ggplot(data = <NAME OF DATASET>, 
       mapping = aes(x = <NAME OF VARIABLE>, y = <NAME OF VARIABLE>)
       ) + 
  geom_boxplot() + 
  labs(x = "<TITLE FOR THE X-AXIS>", 
       y = "<TITLE FOR THE Y-AXIS>")

Note: For horizontally stacked boxplots, the categorical variable should be on the x-axis. For vertically stacked boxplots, the categorical variable should be on the y-axis. 

 
Calculating Summary Statistics for One Numerical Variable and One Categorical Variable

favstats(<NAME OF NUMERICAL VARIABLE> ~ <NAME OF CATEGORICAL VARIABLE>,
         data = <NAME OF DATASET>)

Note: The ~ must be included! 


