---
title: "Week One: Foundations of Statistics"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
library(flair)
library(emo)

library(unvotes)
library(lubridate)
library(DT)
```


Welcome!

In this coursework, you'll get a refresher on the foundational components of 
statistics and data, investigate how statistics is used in your major, and
think critically about the philosophy of statistical inference.

## Course TODOs

`r emo::ji("play button")` Watch Video: 20 min

`r emo::ji("book")` Readings: 60-75 minutes 

`r emo::ji("computer")` Activities: 15 min 

`r emo::ji("check")` Check-ins: 1

`r emo::ji("paper")` Group Discussion Homework: 5 Questions

---

# Data Refresher 

Let's start off by reading a refresher on some of the foundational concepts for 
data! 

`r emo::ji("book")` [Required Reading: *Introduction to Modern Statistics* -- Hellow Data](https://openintro-ims.netlify.app/data-hello.html)

`r emo::ji("book")` [Required Reading: *Introduction to Modern Statistics* -- Study Design](https://openintro-ims.netlify.app/data-design.html)

`r emo::ji("book")` [Required Reading: statistics versus Statistics ](https://critical-inference.com/statistics-vs-statistics/)


`r emo::ji("check")` **Check-in 1:** 

1. What are the different types of variables that can appear in a 
dataset? 

- discrete numerical
- ordinal categorical
- nominal categorical 
- continuous numerical
- discrete categorical

2. Just because a variable has numeric values, does not mean it is a numeric
variable. How can you determine if a variable is numerical versus categorical? 

3. Which of the following statements are true about observational studies and
experiments?

- Experiments randomly assign the explanatory variable
- Observational studies randomly assign the explanatory variable
- Observational studies can make causal statements about the relationship
between the explanatory and response variables
- Experiments can make causal statements about the relationship between the 
explanatory and response variables

4. What are different methods for sampling from a population? 

- simple random sampling 
- stratified random sampling 
- cluster sampling 
- multistage sampling
- convenience sampling

5. What does the "stratification" in a stratified random sample mean? 

</br>

---

# Statistics in Your Field

Now let's investigate how statistics is used in your discipline. 

Find a peer-reviewed article from a journal in your field that performs a 
statistical analysis somewhere in the paper. You **do not** need to be familiar
with the analysis they do. That being said, if you want to find a paper that
uses a statistical method we'll cover in this class, look for a paper that uses
"regression" or "ANOVA." 

The Library is a great resource for this! The following website [https://csu-calpoly.primo.exlibrisgroup.com/discovery/jsearch?vid=01CALS_PSU:01CALS_PSU](https://csu-calpoly.primo.exlibrisgroup.com/discovery/jsearch?vid=01CALS_PSU:01CALS_PSU) has a list of journals by discipline that are available through the
library. 

`r emo::ji("computer")` **Check-in 2:** Upload a PDF of your journal article to 
the Canvas assignment portal. 

</br>

---

# The Philosophy of Statistical Inference 

`r emo::ji("book")` [**Required Reading: Our love of dichotomizing**](https://critical-inference.com/our-love-of-dichotomizing/)

`r emo::ji("book")` [**Required Reading: Science is Human**](https://critical-inference.com/science-is-human/)

`r emo::ji("book")` [**Required Reading: Do We Really Need the S-word?**](https://www.americanscientist.org/article/do-we-really-need-the-s-word)

</br>

`r emo::ji("paper")` **Group Homework Questions**

Throughout the week, your group will be asked to discuss the following 
questions:

- What is **S**tatistics?

- What is the role of Statistics in the progression of science?

- How is dichotomous thinking used in Statistics? What are the benefits and 
drawbacks?

- What does it mean for an analysis to be "objective"? Do you believe 
statistical analyses can be objective? Why or why not?

- What are some synonyms of "significant" that come to your mind?

- What are two issues that you see with using the term "significant" to denote 
"small" p-values? 

</br> 

The reporter in your team will be expected to type up your team's responses to 
these questions. The editor in your team is responsible for submitting the 
completed questions to the Canvas assignment portal. 
