---
title: "Week Eight: Statistical Inference with Hypothesis Tests"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
library(emo)
library(gridExtra)
```


Welcome!

In this week's coursework we are putting all of the pieces we've learned 
together: linear regression, confidence intervals, and p-values! We are going 
to use your knowledge of linear regression to make inferences about the value 
of the population slope. 

We are going to use permutation distributions to obtain other slope statistics
that we could have observed **if the null hypothesis was true**. We are going to
use bootstrap distributions to obtain other slope statistics we could have 
observed from repeated samples from the population. Then, we will use these 
distributions to make inferences about the value of the population slope
($\beta$).


## Learning Outcomes 

By the end of this coursework you should be able to:

<!-- p-values -->
<!-- - describe what a permutation distribution is -->
<!-- - detail how a permutation distribution is made -->
<!-- - describe how permuting assumes the null hypothesis is true -->
<!-- - explain how a permutation distribution relates to a sampling distribution -->
<!-- - use R to:  -->
<!--   * generate a permutation distribution -->
<!--   * visualize the permutation distribution -->
<!--   * calculate the observed statistic -->
<!--   * calculate the resulting p-value -->
<!-- - outline what a p-value represents -->
<!-- - diagnose misinterpretations of a p-value -->

<!-- Confidence intervals -->
<!-- - describe what a bootstrap distribution is -->
<!-- - detail how a bootstrap distribution is made -->
<!-- - explain how a bootstrap distribution relates to a sampling distribution -->
<!-- - use R to: generate a bootstrap distribution, visualize the bootstrap -->
<!-- distribution, & calculate the resulting confidence interval -->
<!-- - outline what a confidence interval is estimating -->
<!-- - diagnose incorrect interpretations of a confidence interval -->



---

## This Week's TODOs

`r emo::ji("book")` Reading: 60 - 75 minutes 

`r emo::ji("video")` Video: 8 minutes

`r emo::ji("paper")` Group Discussion Homework: 9 questions  

`r emo::ji("check")` Check-ins: 2

---

# Hypothesis Tests 

`r emo::ji("book")` [**Required Reading:** *Modern Dive*, Inference for Regression ](https://moderndive.com/10-inference-for-regression.html)

---

The TODO tasks are broken up by the section of the textbook. Each section has
either (1) group concept questions or (2) check-in questions. 

## Section 10.1 - Regression Refresher

1. What parameter are we (typically) interested in for linear regression? What 
symbol do we use to notate this parameter? 

</br>

## Section 10.2 - Interpreting Regression Tables

### `r emo::ji("paper")` Group Concept Questions

2. What does the "standard error" output in the regression table *mean*? i.e. 
What does this value represent? 

3. How is the "statistic" in the regression table found? How does this differ 
from the observed slope? 

4. The null hypothesis for testing a slope is $\beta = 0.$ What does this 
hypothesis *mean* (i.e. what are you testing)?

</br> 

### `r emo::ji("check")` Check-in: Interpreting Regression Tables

<!-- SE -->

<!-- Statistic -->

<!-- p-value -->

<!-- CI -->

</br>

## Section 10.3 - Conditions for Inference for Regression

### `r emo::ji("video")` Independence Violations

Video link: <http://somup.com/crX20q0hwW>

### `r emo::ji("paper")` Group Concept Questions

5. What are three ways that observations could not be independent? 

6. Can we ever *prove* that these conditions for inference are "met"? Why or 
why not?

</br>

### `r emo::ji("check")` Check-in: Conditions for Inference

1. What are the required conditions for linear regression?  

- linear relationship between $x$ and $y$ 
- a random sample was taken
- independence of variables
- independence of observations 
- a large sample was collected
- normality of residuals
- normality of observations
- equal variance of residuals

2. Which of the following would violate the condition of independence?

- repeated observations on the same person 
- collecting a non-random sample
- observations are related in time (temporal correlation)  
- observations are geographically related (spatial correlation)
- observations are biologically related 

3. Which of the following would violate the condition of normality? 

```{r, echo = FALSE}

set.seed(12345)

norm <- tibble(value = rnorm(n = 50)
               )

t <- tibble(value = rt(n = 50, df = 15)
            )

beta1 <- tibble(value = rbeta(n = 50, shape1 = 2, shape2 = 5)
               ) %>% 
  mutate(value = value - mean(value))

beta2 <- tibble(value = rbeta(n = 50, shape1 = 4, shape2 = 10)
            ) %>% 
  mutate(value = value - mean(value))


p1 <- norm %>% 
  ggplot(aes(x = value)) + 
  geom_histogram(binwidth = 0.4) +
  labs(title = "Plot 1")

p2 <- t %>% 
  ggplot(aes(x = value)) + 
  geom_histogram(binwidth = 0.75) +
  labs(title = "Plot 2")

p3 <- beta1 %>% 
  ggplot(aes(x = value)) + 
  geom_histogram(binwidth = 0.07) +
  labs(title = "Plot 3")

p4 <- beta2 %>% 
  ggplot(aes(x = value)) + 
  geom_histogram(binwidth = 0.08) + 
  labs(title = "Plot 4")

grid.arrange(p1, p2, p3, p4, nrow = 2)

```


4. Which of the following would violate the condition of equal variance?

- observations that (when plotted) look like a cloud of points
- observations that look like a funnel
- observations that look like a sine wave

5. What condition for inference are simulation-based methods more resistant to? 

Type in the letter corresponding to the condition: L, I, N, or E. 

</br>

## Section 10.4 - Simulation Based Inference for Regression

### `r emo::ji("paper")` Group Concept Questions

7. How is the p-value in the regression table found? How does this differ from 
simulation-based methods?

8. How is the confidence interval in the regression table found? How does this
differ from simulation-based methods?

9. Why would someone choose to use simulation-based methods over theoretical
($t$-based) methods? 
