---
title: "Introduction to Visualizing and Summarizing Data"
author: "Your Name Here!"
date: "Due: January 24, 2021 at 11:59pm"
output: html_document
---

<!-- The code chunk below is made to make your report look nice. Do not change it! 
-->
```{r, include = FALSE}
knitr::opts_chunk$set(echol = TRUE, 
                      results = "hold", 
                      message = FALSE, 
                      warning = FALSE)
```

## Getting started

### Load packages

In this lab, we will explore and visualize the data using packages housed in the
**tidyverse** suite of packages. 

```{r, include = FALSE}
library(tidyverse)
```

### The data

In this lab we will work with data from the Star Wars API 
(<https://swapi.dev/>). The Star Wars API, or "swapi" (Swah-pee) is the world's
first quantified and programmatically-accessible data source for all the data
from the Star Wars canon universe!

First, we'll view the `starwars` dataframe. Run the code below to load the 
data into your R workspace. 

```{r}
data(starwars)
```

The data set `starwars` that shows up in your workspace is a *data matrix*,
with each row representing an *observation* and each column representing a
*variable*. `R calls this data format a **data frame**, which is a term that
will be used throughout the labs. For this data set, each *observation* is a
Star Wars character.

To view the names of the variables, you can use the `names()` function: 

```{r}
names(starwars)
```

This returns the names of the variables in this data frame.
The **codebook** (description of the variables) can be accessed by pulling up
the help file by typing a `?` before the name of the dataset:

```{r, eval = FALSE}
?starwars
```

Remember that you can use `glimpse` to take a quick peek at your data to
understand its contents better.

```{r}
glimpse(nycflights)
```

1. (a) How large is the `starwars` dataset? (i.e. How many rows and columns
does it have?)

```{r}
# Your code to answer question 1 goes here! 

```

1. (b) Are there categorical variables in the dataset? If so, what are their
names?  
<!-- Your response goes below this line! -->


# Accessing the Levels of a Variable

The `gender` variable refers to the gender expression of the character. 

2. What are the distinct levels / values of the `gender` variable? 

```{r}
# Your code to answer question 2 goes here! 

```


# Creating New Variables 

Suppose we want to use the `species` variable to make a new variable called
`human`. This new variable should have the value of "human" if the `species` of 
the character is human and "not human" otherwise. 

3. Use the `mutate()` and `if_else()` functions to create a new variable named
`human` based on the criteria outlined above. 

```{r}
# Your code to answer question 3 goes here! 

```


# Filtering the Data 

Now, we are interested in creating a new dataset that only contains the 
human characters. 

4, Using the new `human` variable, create a new dataset named
`human_characters` that only contains observations where the value of `human` 
is equal to "human". 

Hint: The `filter()` function would be helpful here! 

```{r}
# Your code to answer question 4 goes here! 

```

# Using the Pipe Operator Instead 

The pipe operator (%>%) allows us to connect our data wrangling steps together. 
Let's get some practice chaining data wrangling operations together by
connecting the previous two steps with the pipe operator.

5. Take your code from problems 3 and 4 and combine them so that the two steps
are connected with a pipe operator. Specifically you should: 

- take the `starwars` data
- create a new `human` variable
- filter the data based on the `human` variable
- save the resulting data as a new dataset named `human_characters`

```{r}
# Your code to answer question 5 goes here! 

```

----

# Summarizing a Categorical Variable

6. Now, using the `human_characters` dataset you made, summarize the number 
of characters from each `homeworld`. 

```{r}
# Your code to answer question 6 goes here! 

```

# Visualizing a Categorical Variable 

7. Create a visualization of the summaries you found in question 6. 

```{r}
# Your code to answer question 7 goes here! 

```

**Bonus:** Add a second categorical variable to your visualization! 
