---
title: "Plotting Numerical Data"
author: "Dr. Allison Theobold"
date: "October __, 2021"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "slide-style.css"]
    nature:
      highlightStyle: solarized-light
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---        

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# R options
options(
  htmltools.dir.version = FALSE,
  tibble.width = 65,
  width = 65
  )

# figure height, width, dpi
knitr::opts_chunk$set(fig.width = 8, 
                      fig.asp = 0.618,
                      out.width = "60%",
                      dpi = 300)

# fontawesome
htmltools::tagList(rmarkdown::html_dependency_font_awesome())

# magick
dev.off <- function(){
  invisible(grDevices::dev.off())
}

# xaringanExtra
library(xaringanExtra)
xaringanExtra::use_panelset()

library(emo)
library(palmerpenguins)
library(tidyverse)
library(flair)
```

```{r set-theme, include = FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color      = "#b76352", # mango
  secondary_color    = "#34605f", # bayberry
  header_font_google = google_font("Raleway"),
  text_font_google   = google_font("Raleway", "300", "300i"),
  code_font_google   = google_font("Source Code Pro"),
  header_color = "#793540", #rhubarb
  white_color = "#F5F5F5", # lightest color
  black_color = "#36454F", # darkest color
  text_font_size = "30px", 
  link_color = "#696969" #grey
)
```

.larger[Color the Points]

.pull-left[
```{r color, echo = FALSE}
ggplot(data = mpg) + 
  geom_point(
    mapping = aes(x = displ,
                  y = hwy,
                  color = cyl)
    )
```
]

.pull-right[
```{r, echo = FALSE}
decorate_chunk(chunk_name = "color", eval = FALSE) %>% 
  flair(pattern = "color = cyl", color = "red")
```

- `r flair("fill", pattern = "color", color = "red")` points based on
the number of **`cyl`**inders the vehicle has
]

---
.larger[Shape]

.pull-left[
```{r shape, echo = FALSE, warning = FALSE,}
ggplot(data = mpg) + 
  geom_point(
    mapping = aes(x = displ,
                  y = hwy,
                  shape = class))
```
]

.pull-right[
```{r, echo = FALSE}
decorate_chunk(chunk_name = "shape", eval = FALSE) %>% 
  flair(pattern = "shape = class", color = "purple")
```

- Uses a different `r flair("shape", pattern = "shape", color = "purple")` points based on **`class`** of vehicles
]

---

.larger[Size]

.pull-left[
```{r size, echo = FALSE, warning = FALSE}
ggplot(data = mpg) + 
  geom_point(
    mapping = aes(x = displ, 
                  y = hwy, 
                  size = cyl))
```
]

.pull-right[
```{r, echo = FALSE}
decorate_chunk(chunk_name = "size", eval = FALSE) %>% 
  flair(pattern = "size = cyl", color = "red")
```

- Uses a different `r flair("size", pattern = "size", color = "red")` points based on **`ncyl`** of the vehicle
]

---

.larger["Plain" Aesthetics]

.pull-left[
.large[
Assign a single `color`, `shape`, or `size` to **all** points 
]
```{r, eval = FALSE}
ggplot(data = mpg) + 
  geom_bar(mapping = aes(y = manufacturer), 
           color = "tomato")
```
]

.pull-right[

```{r, echo = FALSE, fig.width = 6, fig.height = 6}
ggplot(data = mpg) + 
  geom_bar(mapping = aes(y = manufacturer), 
           fill = "tomato")
```
]