---
title: "Lab 2 - Grading Guide"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

__1. (a) How large is the `nycflights` dataset? (i.e. How many rows and columns does it have?)__
<!-- Your response goes below this line! -->



__1. (b) Are there numerical variables in the dataset? If so, what are their names?__  
<!-- Your response goes below this line! -->


__2. Create a histogram of the `dep_delay` variable from the `nycflights` data__ 

```{r histogram}
# Your code for question 2 goes here! 


```


__3.  (a) Make two other histograms, one with a `binwidth` of 15 and one with a `binwidth` of 150.__

```{r binwidth}
# Your  code for question 3 goes here! 


```

__3. (b) How do these three histograms compare? Are features revealed in one that are obscured in another?__
<!-- Your answer goes below this line -- do not remove this!  -->


__4. Create a new dataframe named `sfo_flights` that contains only flights whose destination was San Francisco.__ 

```{r}
# Your code for question 4 goes here! 


```

__5.  Use your `sfo_flights` dataset to find the number of flights in July that arrived early.__ 

```{r}
## Code for exercise 5 here! 


```

__6. Find the average departure delays for each month.__ 
**Bonus:** Arrange these average delays in ascending order. 
**Hint:** The `arrange()` function is helpful for arranging numbers!

```{r}
## Code for exercise 6 goes here! 


```

__7. Now, rather than calculating means plot the distribution of departure delays, faceted by the month of departure.__

You choose the type of plot you believe is appropriate for visualizing the **distribution** of departure delays. 

```{r}
## Code for exercise 7 goes here! 


```

__8. What information can you obtain from the visualization that you could not from the data summaries?__
<!-- Your response to this question goes below this line!  -->
