---
title: "Lab 2 - Grading Guide"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)

library(tidyverse)
library(openintro)
```

## 1 (a)
__How large is the `nycflights` dataset? (i.e. How many rows and columns does it have?)__
<!-- Your response goes below this line! -->

**4**:

- Must have code that gives size of dataset, with a preference for `glimpse()`
- Size of 32,735 rows and 16 columns

**Redo**:

- If no code is present
- If they flip the rows and columns
- If they do not provide size of data


## 1 (b)
__Are there numerical variables in the dataset? If so, what are their names?__  
<!-- Your response goes below this line! -->

**4**: 

- Lists all variables with an `int` or `dbl` data type: 
  * year, month, day, dep_time, dep_delay, flight, air_time, distance, 
  hour, minute
  
**3**:

- If misses 1-2 variables

**Redo**:

- If misses more than 2 variables
- If includes categorical variables (labeled `chr`)

## 2 

__Create a histogram of the `dep_delay` variable from the `nycflights` data__ 

**4**: Code should look like 

```{r histogram}
nycflights %>% 
  ggplot(aes(x = dep_delay)) + 
  geom_histogram()

```

**Redo**: If maps `dep_delay` to y-axis instead of x-axis

## 3 (a)

__Make two other histograms, one with a `binwidth` of 15 and one with a `binwidth` of 150.__

**4**: Code should look like: 

```{r binwidth}
nycflights %>% 
  ggplot(aes(x = dep_delay)) + 
  geom_histogram(binwidth = 15)

nycflights %>% 
  ggplot(aes(x = dep_delay)) + 
  geom_histogram(binwidth = 150)

```

**Redo**: If code does not use correct `binwidth`s

## 3 (b)

__How do these three histograms compare? Are features revealed in one that are obscured in another?__
<!-- Your answer goes below this line -- do not remove this!  -->

**4**: 

- Discusses how wider bins make it hard to see where there are small
differences and / or how smaller bins make it easier to see differences

**3**:

- If description seems to be referring to the binwidth and how it makes wider /
narrower bins, but is not very clear

**Redo**:

- Only talks about how the y-axis scale changes


## 4 

__Create a new dataframe named `sfo_flights` that contains only flights whose destination was San Francisco.__ 

**4**: Code should look like:

```{r}

sfo_flights <- nycflights %>% 
  filter(dest == "SFO")

```

**Redo**: If they use the wrong destination in their `filter()` 

## 5 
__Use your `sfo_flights` dataset to find the number of flights in July that arrived early.__ 

**4**: Code should look like:

```{r}
sfo_flights %>% 
  filter(month == 7, 
         arr_delay < 0) %>% 
  count()
```

- The final call to `count()` could also be one of the following functions:
  * `dim()`
  * `glimpse()`
  
**Redo**:

- If use `nycflights` dataset instead of `sfo_flights`
- - If the final function doesn't output the size of the dataset (e.g., 
`summary()`)


## 6 

__Find the average departure delays for each month.__ 
**Bonus:** Arrange these average delays in ascending order. 
**Hint:** The `arrange()` function is helpful for arranging numbers!

**4**: Code should look like:

```{r}
nycflights %>% 
  group_by(month) %>% 
  summarize(mean_dd = mean(dep_delay)) %>% 
  arrange(mean_dd)


```

- I'd give full credit for using either the `nycflights` dataset or the 
`sfo_flights` dataset

- I'd write a comment if they didn't name their summary, suggesting they do 
since it makes arranging much easier!

- I'd write a comment if they don't use a %>% for the first line (e.g., `group_by(nycflights, month)`), suggesting they start with a pipe operator
instead. 

**Redo**: 

- If their code doesn't follow steps 1 - 3 from above. 

## 7 

__Now, rather than calculating means plot the distribution of departure delays, faceted by the month of departure.__

You choose the type of plot you believe is appropriate for visualizing the **distribution** of departure delays. 

**4**: Code should look like:

```{r}
nycflights %>% 
  ggplot(mapping = aes(x = dep_delay)) + 
  geom_histogram() + 
  facet_wrap(~ month)


```

- I'd give full credit for using either the `nycflights` dataset or the 
`sfo_flights` dataset

- Acceptable `geom`s:
  * `geom_histogram()`
  * `geom_dotplot()` (though not great)
  * `geom_density()`

**Redo**: 

- Doesn't use facets
- Uses `geom_boxplot()`

## 8 

__What information can you obtain from the visualization that you could not from the data summaries?__

**4**: Answer should address *all* of the following

- visualization allows for you to see skew
- visualization allows for you to see mode / peak

**3**: Answer addresses either skew or modes, but not both

**Redo**: 

- Response doesn't make much sense
- Response doesn't address the shape of the distribution

