---
title: "Week Eight: Statistical Inference with Hypothesis Tests"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
library(emo)
```


Welcome!

In this week's coursework we are **finally** talking about p-values! What we
learned last week should have helped make the connection between a sampling 
distribution and a bootstrap (resampling) distribution. Hopefully, you
understand that we create confidence intervals based on the values of the 
statistic we saw in other samples. 

This week we are going to connect these ideas to the framework of hypothesis 
testing. Hypothesis testing requires an additional component we didn't see 
last week, the null hypothesis. We will learn how we integrate the null
hypothesis in our resampling procedure to create a sampling distribution that 
could have happened if the null hypothesis was true. We will use this
distribution to compare what we saw in our data and decide if what we saw was
"rare" or not. 

## Learning Outcomes 

By the end of this coursework you should be able to:

- describe what a permutation distribution is
- detail how a permutation distribution is made
- describe how permuting assumes the null hypothesis is true
- explain how a permutation distribution relates to a sampling distribution
- use R to: 
  * generate a permutation distribution
  * visualize the permutation distribution
  * calculate the observed statistic
  * calculate the resulting p-value
- outline what a p-value represents
- diagnose misinterpretations of a p-value

---

## This Week's TODOs

`r emo::ji("book")` Reading: 60 - 75 minutes 

<!-- `r emo::ji("computer")` Tutorial: 45 - 60 minutes -->

`r emo::ji("paper")` Group Discussion Homework: 7 questions  

`r emo::ji("check")` Check-ins: 2

---

# Hypothesis Tests 

`r emo::ji("book")` [**Required Reading:** *Modern Dive*, Hypothesis Testing](https://moderndive.com/9-hypothesis-testing.html)

---

This week, I'm breaking up the TODO tasks by section of the textbook. Each 
section has either (1) group concept questions, (2) an R tutorial, or (3) 
check-in questions. 

## Section 9.1 

# `r emo::ji("paper")` Group Discussion Question

1. Why is the distribution of shuffled differences in promotion rates (in Figure
9.6) centered at 0?

---

## Section 9.2

# `r emo::ji("check")` Check-in: Understanding Hypothesis Tests 

1. Which of the following is **always** true for hypothesis statements? 

- They are about sample statistics.
- They are about population parameters.
- They are about differences in groups.
- They are about relationships between variables.

2. Which of the following are true about a null distribution?

- It is a distribution of statistics. 
- It is a distribution of observations. 
- The values on the distribution represent what might have happened if the 
null hypothesis was true. 
- The values on the distribution represent what might have happened if the 
alternative hypothesis was true. 

3. Which of the following are true about a p-value?

- It is a probability. 
- It quantifies how "surprising" our data are. 
- It is calculated assuming the null hypothesis is true. 
- It is calculated assuming the alternative hypothesis is true. 
- It compares the observed statistic to a distribution of values that could 
have happened if the null was true. 
- It is the probability the null hypothesis is true.

---

## Section 9.3

# `r emo::ji("paper")` Group Discussion Questions

2. When creating a null distribution, what are the sequence of functions you 
use? 

3. Why is the `hypothesize()` function used to make a null distribution but not 
for a bootstrap distribution?

4. What does the `null = "independence"` input in `hypothesize()` mean? What is 
it assuming about the variables declared in the `specify()` step?

5. A null distribution is created by resampling the original data
*without replacement*. How is this resampling done? How does this connect to 
the null hypothesis? 

<!-- </br>  -->

<!-- # `r emo::ji("computer")` Tutorial -->

<!-- Work through this short tutorial giving you experience coding the process  -->
<!-- discussed in this section.  -->

<!-- Tutorial: <link> -->

---


## Section 9.4

# `r emo::ji("check")` Check-in: Interpreting Hypothesis Tests 

1. Which of the following are true about a small p-value?  

- The sample statistic was large. 
- The sample size was large. 
- The sample statistic is unlikely to have happened by chance. 
- The sample statistic is unlikely to have happened if the null hypothesis was true. 

2. If you obtain a large p-value, what can you conclude about your hypotheses?

- The null hypothesis is true. 
- The alternative hypothesis is true. 
- We cannot say the null hypothesis is false. 
- We cannot say the alternative hypothesis is false. 

3. If you obtained a small p-value (e.g., 0.02), what could you say about your 
confidence interval? 

- It would contain the true population parameter. 
- It would contain the sample statistic. 
- It would not contain the null hypothesized value. 
- It would contain the null hypothesized value. 

4. Match each procedure to the question it addresses. 

__confidence intervals__    _hypothesis tests_ 

- Is the value of the parameter different from a specified quantity?
- What are plausible values for the population parameter?

5. If the probability of a Type I error goes down, what can you say about the
probability of a Type II error?

- The probability of a Type II error goes down. 
- The probability of a Type II error goes up. 
- The probability of a Type II error stays the same. 

---

# `r emo::ji("paper")` Group Discussion Questions

6. What is the relationship between $\alpha$ and a Type I error? Why is one 
benefit and one cost of choosing a smaller value of $\alpha$ versus a larger
value of $\alpha$?

---

## Section 9.6

This section references many articles addressing the "p-value" crisis. I would
like for **everyone** to read the ASA's *original* statement on p-values and 
statistical significance. It is crazy that this statement is over 5 years old, 
and a lot has happened since. But, I think this statement captures some of the 
*critical* misconceptions and misinterpretations of p-values. 

[ASA Statement on Statistical Significance and p-values](https://www.amstat.org/asa/files/pdfs/p-valuestatement.pdf)

# `r emo::ji("paper")` Group Discussion Questions

7. Pick one of the misinterpretations of p-values and talk about *why* it is a
misinterpretation. 

---

## Think Out Loud

Look at the research article you found in Week 1. Turn to the analysis section 
and hunt for a statement with a p-value. Carefully read this statement and the
surrounding mathematical information (e.g., p < 0.001). 

In your recording I expect you to discuss:

- What p-value was obtained by the authors (for this analysis)?
- What p-value threshold was used?
- What was said about why that threshold was chosen?
- How did the authors present the results of their statistical analysis?
  * What conclusions were reached?
  * Was the p-value interpreted in the context of the study?
  * Was the observed statistic discussed? 
  
