---
title: "Untitled"
format: html
editor: visual
---

One of the variables refers to the carrier (i.e. airline) of the flight, which
is coded according to the following system.

-   `carrier`: Two letter carrier abbreviation.

    -   `9E`: Endeavor Air Inc.
    -   `AA`: American Airlines Inc.
    -   `AS`: Alaska Airlines Inc.
    -   `B6`: JetBlue Airways
    -   `DL`: Delta Air Lines Inc.
    -   `EV`: ExpressJet Airlines Inc.
    -   `F9`: Frontier Airlines Inc.
    -   `FL`: AirTran Airways Corporation
    -   `HA`: Hawaiian Airlines Inc.
    -   `MQ`: Envoy Air
    -   `OO`: SkyWest Airlines Inc.
    -   `UA`: United Air Lines Inc.
    -   `US`: US Airways Inc.
    -   `VX`: Virgin America
    -   `WN`: Southwest Airlines Co.
    -   `YV`: Mesa Airlines Inc.

## Grouped Summaries 

Another useful technique is quickly calculating summary statistics for various
groups in your data frame. For example, we can modify the above command using
the `group_by()` function to get the same summary stats for each origin airport:

```{r}
sfo_feb_flights %>%
  group_by(origin) %>%
  summarise(median_dd = median(dep_delay), 
            iqr_dd = IQR(dep_delay), 
            n_flights = n())
```

Here, we first grouped the data by `origin` and __then__ calculated the summary
statistics.


**Question 6** -- Find the average departure delays for each month. 
**Bonus:** Arrange these average delays in ascending order. 
**Hint:** The `arrange()` function is helpful for arranging numbers!

```{r}
## Code for exercise 6 goes here! 


```
