---
title: "Week Three: Summarizing & Visualizing Numerical Data"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
library(flair)
library(emo)

library(unvotes)
library(lubridate)
library(DT)
```


Welcome!

In this week's coursework we are going to continuing exploring data, through 
data summaries and visualizations, focusing specifically on **numerical**
variables. We will be using the **dplyr** package in `R` to wrangle our data 
and the **ggplot2** package to created data visualizations. 


## Learning Outcomes 

By the end of this coursework you should be able to:

- outline the differences between numerical and categorical variables
- describe what type of summary statistic is appropriate for a given  
distribution of a numerical variable  
- discuss when it is / is not appropriate to summarize a variable with a mean  
- create visualizations of one and two numerical variables  
- discuss the benefits and shortcomings of different visualizations 


## This Week's TODOs

`r emo::ji("book")` Reading: 30-45 minutes 

`r emo::ji("computer")` Tutorials: 60-75 minutes 

`r emo::ji("check")` Check-ins: 2 (submit on Canvas)

`r emo::ji("scroll")` Group Discussion Homework: __ Questions

---

# Exploring Numerical Data 

`r emo::ji("book")` [**Required Reading:** Exploring Numerical Data](https://openintro-ims.netlify.app/explore-numerical.html)

`r emo::ji("computer")` [**Required Tutorial:** Summarizing with Statistics](https://openintro.shinyapps.io/ims-02-explore-03/)


`r emo::ji("computer")` [**Required Tutorial:** Exploring Numerical Data](https://openintro.shinyapps.io/ims-02-explore-02/)


If you are feeling like you would like to see additional materials on
wrangling data and creating data visualizations in `R`, the *Modern Dive* 
textbook has two exellent chapters you can reference. 

`r emo::ji("book")` [**Supplementary Reading:** Data Visualization](https://moderndive.com/2-viz.html)

`r emo::ji("book")` [**Supplementary Reading:** Data Visualization](https://moderndive.com/3-wrangling.html)

---

`r emo::ji("check")` **Check-in: Summarizing with Statistics**

---

`r emo::ji("check")` **Check-in: Visualizing Numerical Data**


1. What aesthetics are being used in the following plot?  

Hint: Think about what goes inside of the `aes()` function and what does not. 

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "70%", fig.align = 'center'}
un_votes %>%
  filter(country %in% c("United States", "Turkey")) %>%
  inner_join(un_roll_calls, by = "rcid") %>%
  inner_join(un_roll_call_issues, by = "rcid") %>%
  mutate(issue = ifelse(issue == "Nuclear weapons and nuclear material",
                        "Nuclear weapons and materials", issue)) %>%
  group_by(country, year = year(date), issue) %>%
  summarize(
    votes = n(),
    percent_yes = mean(vote == "yes")
    ) %>%
  filter(votes > 5) %>%  # only use records where there are more than 5 votes
  ggplot(mapping = aes(x = year, y = percent_yes, color = country)) +
    geom_point() +
    geom_smooth(method = "loess", se = FALSE) +
    facet_wrap(~ issue) +
    labs(
      title = "Percentage of 'Yes' votes in the UN General Assembly",
      subtitle = "1946 to 2020",
      y = "% Yes",
      x = "Year",
      color = "Country"
    )
```

- `x` axis
- `y` axis
- `color` 
- `facet`s
- `fill`
- `point`s
- `line`s
