---
title: "Week One: Why Statistics?"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
library(flair)
library(emo)

library(unvotes)
library(lubridate)
library(DT)
```


Welcome!

In this coursework, you'll get a refresher on the foundational components of 
statistics and data, investigate how statistics is used in your major, and
think critically about the philosophy of statistical inference.

## Course TODOs

`r emo::ji("play button")` Watch Video: 20 min

`r emo::ji("book")` Readings: 60-75 minutes 

`r emo::ji("computer")` Activities: 15 min 

`r emo::ji("check")` Check-ins: 1

`r emo::ji("paper")` Group Discussion Homework: 5 Questions

---

`r emo::ji("check")` **Check-in 1:**


1. Describe what you see in the plot below. 

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "70%", fig.align = 'center'}
un_votes %>%
  filter(country %in% c("United States", "Turkey")) %>%
  inner_join(un_roll_calls, by = "rcid") %>%
  inner_join(un_roll_call_issues, by = "rcid") %>%
  mutate(issue = ifelse(issue == "Nuclear weapons and nuclear material",
                        "Nuclear weapons and materials", issue)) %>%
  group_by(country, year = year(date), issue) %>%
  summarize(
    votes = n(),
    percent_yes = mean(vote == "yes")
    ) %>%
  filter(votes > 5) %>%  # only use records where there are more than 5 votes
  ggplot(mapping = aes(x = year, y = percent_yes, color = country)) +
    geom_point() +
    geom_smooth(method = "loess", se = FALSE) +
    facet_wrap(~ issue) +
    labs(
      title = "Percentage of 'Yes' votes in the UN General Assembly",
      subtitle = "1946 to 2020",
      y = "% Yes",
      x = "Year",
      color = "Country"
    )
```