---
title: "Introduction to Sampling"
date: "February 9, 2021"
output: ioslides_presentation
---

<style type="text/css">
slides > slide:not(.nobackground):after {
  content: '';
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      echo = FALSE, 
                      fig.width = 4, 
                      fig.height = 4, 
                      message = FALSE)
library(tidyverse)
library(openintro)
library(gridExtra)
library(moderndive)
library(infer)
```

## Sampling Concepts 

- A __population__ is a collection of individuals or observations we are interested in. 
  * A __population parameter__ is a numerical summary quantity about the population that is unknown, but you wish you knew.

</br> 

- __Sampling__ is the act of collecting a sample from the population when we don’t have the means to perform a census.
  * A __point estimate__ is a summary statistic computed from a sample that estimates an unknown population parameter. 

## Population Parameter 

To investigate the relationship between course evaluation score and the 
professor's age, we would carry out a simple linear regression. 

- What would be the population parameter we are interested in? Do we know its value? 

- What would be our point estimate? What is its purpose? 

```{r, echo = FALSE, fig.align='center', fig.width=4, fig.height=2.75}
evals %>% 
  ggplot(aes(x = age, y = score)) + 
  geom_jitter() + 
  geom_smooth(method = "lm")
```

## Sampling Procedures 

- A __census__ is an exhaustive enumeration or counting of all individuals. 

- A __representative sample__ is a sample that looks roughly like the population. 

- A __random sample__ is a sample where every observation has the same probability of selection. 

- A __biased sample__ is a sample where certain individuals or observations in a population have a higher chance of being included than others. 

- __Generalizability__ is when the results from a sample can be _inferred_ to the population. 

