---
title: 'Stat 313: Project 2'
output:
  pdf_document: default
  html_document: default
---

# Write-Up Guidelines

You will be summarizing your results in a (guided) written report following the
__Project Writing Guidelines__ posted on Canvas. The results __must__ be 
written in the RMarkdown template provided. You can include all pertinent plots
inline in the typed document, rather than in the appendix. Your group will be 
submitting both the RMarkdown file used to generate your report and the knitted 
HTML file to Canvas by __Sunday, November 22 at 11:59pm__. 

Begin by downloading the RMarkdown and data files from Canvas and saving them on
your computer in the __same__ folder. Then open the RMarkdown file in RStudio 
and edit the first chunk of code reading the data in. Once you've modified the 
code to read in the data, click "Knit." If you encounter an error, 
make sure that the data are located in the same folder as the RMarkdown file! 

# Data Background 

Tager, Weiss, Rosner, and Speizer (1979), and Tager, Weiss, Muno, Rosner, and
Speizer (1983) reported analyses of a study aimed at assessing children's
pulmonary function in the absence or presence of smoking cigarettes, as well as
exposure to passive smoke from at least one parent. These papers represent some
of the earliest attempts at systematic documentation regarding obvious signs of
reduced pulmonary function from smoking and from exposure to second-hand smoke.

Rosner (1999) presented another investigation of subjects' respiratory function, 
and its relationship with smoking. The data are "cross
sectional," measuring participants at one point in time, rather than following 
them throughout their adolescence. During the study, researchers measured 
participant's height (in cm) and their  forced expiratory volume (FEV; in liters),
which is, essentially, the amount of air an individual can exhale in the first
second of a forceful breath. Following these measurements, participants
completed a survey about:  

- their age (coded as the stage of development)
- their sex (male, female) 
- whether they smoked (non-smoker, smoker)

Our primary question of interest is, controlling for age, whether the pulmonary
function of smokers differs from non-smokers. 

\newpage


# Introduction (5 pts)

- (2 pts) Give a brief background of the research problem and how the data were
collected.

- (3 pts) Clearly outline the question(s) of interest you will address with the
statistical analysis. The more specific you define the question of interest
here, the easier the rest of the analysis and report will be. The research
questions should be as specific as possible! Your 
_Summary of Statistical Findings_ should directly answer the question you
pose here.

\vspace{0.2cm}

# Statistical Methods (20 pts)

This section should lay out the steps, decisions, and logic leading to the
statistical model you will use to answer the research question of interest.

- (5 pt) Describe the response and explanatory variables.   
  * Provide a table summarizing the number of observations for every combination
  of smoking and sex (i.e. non-smoker & female, non-smoker & male, 
  smoker & female, smoker & male).
  * Provide a summary of the ages for male and female participants. 

```{r, echo = FALSE, eval = FALSE}
lungs %>% 
  group_by(smoke, sex) %>% 
  count()
```

\vspace{0.2cm}

- (3 pts) Provide a data visualization of the relationship between FEV and 
smoking status.   
  * Discuss the relationship seen in the plot, making direct reference to the
  plot. 
  * Additionally, discuss whether this relationship is expected __and__
  variables that might be confounding this relationship. 

```{r, echo = FALSE, eval = FALSE}
lungs %>% 
  ggplot(aes(y = FEV, x = smoke)) + 
  geom_violin()
```


\vspace{0.2cm}

- (3 pts) Provide an additional data visualization of the relationship between
FEV and smoking status, accounting for the confounding variable you outlined
above. 
  * **Note:** Keep in mind the different methods for adding a second categorical 
explanatory to your plot (colors or facets)! 

```{r, echo = FALSE, eval = FALSE}
lungs %>% 
  ggplot(aes(y = FEV, x = smoke)) + 
  geom_violin(aes(fill = age))

lungs %>% 
  ggplot(aes(y = FEV, x = smoke)) + 
  geom_violin() +
  facet_wrap(~age)

```

\vspace{0.2cm}

- (5 pts) Describe the appropriate statistical method you will use to answer
the question of interest that you stated previously. 
  * Be specific about why the method being used are appropriate for the
  investigation at hand (e.g. types of variables). 
  * Keep in mind that you should include the confounding variable you outlined 
  above in your statistical model! 
  * Also keep in mind that, like multiple linear regression, there are two
  different ways we can include two categorical variables in our model. 

\vspace{0.2cm}

- (4 pts) Check all model conditions of the statistical method you used. 
  * Describe what each condition is __in the context of these data__. 
  * Reference and include appropriate plots necessary for checking the model
  conditions. 
  * Justify your conclusions regarding the conditions! 

\newpage  

# Summary of Statistical Findings (10 pts)

In this section you will write up your findings for the question of
interest.

- (5 pts) What is your conclusion for the questions of interest? Namely, 
"After controlling for age, does the pulmonary function of smokers differ from
non-smokers?".  
  * Base your conclusion on the visualizations you created and the two-way ANOVA 
  model you fit to these data. 
  * When communicating the results of the statistical test make certain you
  include the value of the test statistic and the distribution it follows under
  the null hypothesis. 
  * Arrive on a conclusion based on the strength of evidence provided by the
  p-value. 
  * If you make a rejection decision, you __must__ specify an $\alpha$ !
  * Do not use the word “significant”! 

_Example for an ANOVA: There is strong evidence that at least one credit rating has a different mean income (p-value = 0.027, F-stat = 3.067 on an F(3, 1616) distribution)._ 

\vspace{0.2cm}
 
- (5 pts) Interpret __in the context of the data__ the p-value associated with 
`smoke` in your ANOVA table. 
  * Keep in mind what the p-value is *conditional* on! 

\vspace{0.2cm}

# Survey Design (5 pts) 

The survey associated with these data specifically asked the study participants 
if they were smokers. 

- What issues do you see with this type of survey question? 
- How would these issues be reflected in the data from this study? 

# Scope of Inference (10 pts)

Write a brief Scope of Inference statement. Specifically, answer these two
questions and comment on their implications:

- (2 pts) Were the observations randomly selected from some larger population?
Based on the sampling method used, what larger population can you infer the
results to? 

- (2 pts) Was the explanatory variable randomly assigned to observations? Based
on the study design, are cause-and-effect statements justified?

Make sure you write the scope of inference __in the context of these data__, not
just generic statements!

\vspace{0.2cm}

# Project Presentation (3 pts) 

- (1 pts) Your report should not have any spelling errors! To check for spelling 
errors in RStudio, click the green check mark button next to the "Knit" button. 

- (2 pts) Your report should look as neat and professional as possible. Make 
sure that your figures don't end up in the middle of your paragraphs, and that 
your sections have headings. 

  * Your plots should be included in the section in which they are discussed! 
  * The output from your statistical model should look as nice as possible. 
  Output from `summary()` looks terrible in a statistical report! I would 
  highly recommend using the `tidy()` function from the **broom** package to 
  tidy up your ANOVA table. 


# Group Evaluation (3 pts)

Each member of your group will fill out a group evaluation form detailing each 
member's contributions, cooperation, communication, and participation.  

- It is not expected that every group member is an expert on these topics. 
- Rather, it is expected that every group member articulates what they are and
are not comfortable contributing.  
- Every member of the group can (and should) contribute to proof reading your 
final report!

