---
title: "Small Group Collaboration - Session 2"
date: "October 1, 2020"
output: ioslides_presentation
---

<style type="text/css">
slides > slide:not(.nobackground):after {
  content: '';
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
library(tidyverse)
library(openintro)
```

## NYC Flights 

```{r, echo = FALSE}
nycflights %>% 
  sample_n(size = 6) %>% 
  glimpse()
```

## Research Questions & Data Wrangling
<center>
Passengers are interested in which month they could expect to have the highest
average delay departing from an NYC airport?

</br>

__How would you wrangle the data to address this question?__ 
</center>
</br>
<!-- First, calculate monthly averages for departure delays.  -->
<!-- - group_by months  -->
<!-- - summarise mean departure delays -->
<!-- Then, you could to arrange these average delays in descending order -->

## Research Questions & Data Wrangling Continued 

- Airlines are interested in the arrival delays for flights headed to San
Francisco (SFO) in the winter (December, January, February). 
<center>
__How would you filter the data to satisfy these criteria?__
</center>
</br>

- For these flights, they are interested in which carrier has the most variable
arrival delays. 
<center>
__How would you address this question?__ 
</center>
<!-- Calculate the median and interquartile range for `arr_delays` of flights in in -->
<!-- the data frame, grouped by carrier. -->

## Research Questions & Data Visualizations

- What is the distribution of departure delays? 

<center>
__What type of visualization would you create?__
</center>
</br>

- How does the distribution of departure delays differ by carrier? 
<center>
__How would you modify your previous visualization?__
</center>
</br>

- How does each carrier's departure delays differ by month?
<center>
__How would you modify your previous visualization?__
</center>

## Effective Data Visualization - Histogram

<div class="columns-2">

```{r, message = FALSE, fig.align = 'center', fig.height = 4, fig.width = 4}
nycflights %>% 
  ggplot(aes(x = arr_delay)) + 
  geom_histogram()
```

</br>
</br>
</br>
</br>
</br>

__How would you change the histogram to more effectively communicate the 
distribution of departure delays?__ 

## Effective Data Visualization - Boxplot

<div class="columns-2">

```{r, message = FALSE, fig.align = 'center', fig.height = 4, fig.width = 4}
nycflights %>% 
  ggplot(aes(y = arr_delay, x = origin)) + 
  geom_boxplot()
```

</br>
</br>
</br>
</br>
</br>

__How would you change the side-by-side boxplots to more effectively communicate
how the distribution of departure delays differs by origin?__ 

## Effective Data Visualization - Scatterplot

```{r, eval = FALSE}
nycflights %>% 
  ggplot(aes(x = distance, y = air_time, color = hour)) + 
  geom_point()
```

<div class="columns-2">

```{r, message = FALSE, fig.align = 'center', fig.height = 3, fig.width = 4, echo = FALSE}
nycflights %>% 
  ggplot(aes(x = distance, y = air_time, color = hour)) + 
  geom_point()
```

</br>

__How would you change the scatterplot to more effectively communicate how the 
relationship between distance and air time differs by the hour the flight 
left the airport?__ 


