---
title: "Week Five: Statistical Inference, Starting with Confidence Intervals"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
library(gapminder)
library(flair)
library(emo)
library(openintro)
library(moderndive)
library(skimr)
library(gridExtra)
```


Welcome!

In this week's coursework we are diving into the land of statistical inference. 
We are going on a bit of a "backwards" journey, starting with confidence 
intervals instead of hypothesis tests. I am doing this for a reason! I want for
you to leave this week with the understanding of how the sampling distributions
from last week connect to the distributions we use to make our confidence 
intervals. 


## Learning Outcomes 

By the end of this coursework you should be able to:



## This Week's TODOs

`r emo::ji("book")` Reading: 60 - 75 minutes 

`r emo::ji("paper")` Group Discussion Homework: __ questions  

`r emo::ji("check")` Check-ins: __ 

---

# Confidence Intervals  

`r emo::ji("book")` [**Required Reading:** *Modern Dive*, Confidence Intervals](https://moderndive.com/8-confidence-intervals.html)

---

# Group Concept Questions

__Question 1__

What does one dot on a bootstrap distribution represent?

</br>

__Question 2__ 

When we resample with replacement from our original sample what are we assuming
about our original sample?

</br>

__Question 3__ 

Why do we resample *with replacement* when creating a bootstrap distribution? 
Why can an observation be selected more than once?  

</br>

__Question 5__

What is the relationship between a bootstrap distribution and a sampling 
distribution?

</br>

__Question 6__

In a bootstrap distribution we resample with replacement from our observed data
the same number of times as samples in the original data. If we were to change
our sample size in the process of creating a bootstrap distribution, what would
happen? 

</br>

__Question 7__

How is a bootstrap distribution used to find a confidence interval?

</br>

__Question 8__

What is one similarity and one difference between using a bootstrap 
distribution to find a confidence interval versus using a t-distribution? 

<!-- Sim: -->
<!-- Both approximate the true sampling distribution -->

<!-- Diff: -->
<!-- Bootstraps resample to approximate the distribution, t-distributions use  -->
<!-- mathematical properties to approximate the distribution -->


---

# Check-in: 1


# Check-in: 2


---

# Think Out Loud Recording 

<!-- Something about sampling variability and sampling distributions -->
