---
title: "Week Five: Statistical Inference, Starting with Confidence Intervals"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
library(gapminder)
library(flair)
library(emo)
library(openintro)
library(moderndive)
library(skimr)
library(gridExtra)
```


Welcome!

In this week's coursework we are diving into the land of statistical inference. 
We are going on a bit of a "backwards" journey, starting with confidence 
intervals instead of hypothesis tests. I am doing this for a reason! I want for
you to leave this week with the understanding of how the sampling distributions
from last week connect to the distributions we use to make our confidence 
intervals. 


## Learning Outcomes 

By the end of this coursework you should be able to:

- describe what a bootstrap distribution is
- detail how a bootstrap distribution is made
- explain how a bootstrap distribution relates to a sampling distribution
- use R to: generate a bootstrap distribution, visualize the bootstrap
distribution, & calculate the resulting confidence interval
- outline what a confidence interval is estimating
- diagnose incorrect interpretations of a confidence interval


## This Week's TODOs

`r emo::ji("book")` Reading: 60 - 75 minutes 

`r emo::ji("computer")` Tutorial: 45 - 60 minutes

`r emo::ji("paper")` Group Discussion Homework: 10 questions  

`r emo::ji("check")` Check-ins: 2

---

# Confidence Intervals  

`r emo::ji("book")` [**Required Reading:** *Modern Dive*, Confidence Intervals](https://moderndive.com/8-confidence-intervals.html)

---

This week, I'm breaking up the TODO tasks by section of the textbook. Each 
section has either (1) group concept questions, (2) an R tutorial, or (3) 
check-in questions. 

## Section 8.1

### `r emo::ji("paper")` Group Concept Questions

**_Question 1_**

Why do we resample *with replacement* when creating a bootstrap distribution? 

**_Question 2_** 

When we resample with replacement from our original sample what are we assuming
about our original sample?

**_Question 3_** 

What does one dot on a bootstrap distribution represent?

**_Question 4_**

What is the relationship between a bootstrap distribution and a sampling 
distribution?

**_Question 5_**

In Section 8.1.2, you see a distribution of the years of pennies for one sample
and one resample. Then, in Section 8.1.3, you see a distribution of the means 
of 35 different samples. 

Why is the bootstrap distribution nearly bell shaped and symmetric when the
distribution of the samples were left skewed? 

---

## Section 8.2

### R Tutorial `r emo::ji("computer")` -- Currently experiencing issues, check back soon!

Do __only__ Sections 1 and 2! <https://rstudio.csm.calpoly.edu/infer_CI_tutorial/>

</br>

### `r emo::ji("paper")` Group Concept Questions

**_Question 6_**

In a bootstrap distribution we resample with replacement from our observed data
the same number of times as samples in the original data. If we were to change
our sample size in the process of creating a bootstrap distribution, what would
happen? 

**_Question 7_**

Where is a bootstrap distribution always centered? 

---

## Section 8.3

### `r emo::ji("check")` Check-in: Creating Confidence Intervals

1. Match each item to it's respective analogy. 

- point estimate 
- confidence interval
- fishing with a spear
- fishing with a net

2. To create a 95% confidence interval using the *percentile method*, what 
percentiles of the bootstrap distribution do you need to calculate?

- 0th
- 2.5th
- 5th
- 90th
- 95th
- 97.5th

3. To create a 95% confidence interval using the *standard error* method, what
standard error do you use?

- 1.96
- the sample standard deviation
- a resample standard deviation
- the bootstrap distribution standard deviation

### `r emo::ji("paper")` Group Concept Questions

**_Question 8_**

How is a bootstrap distribution used to find a confidence interval?

**_Question 9_**

What quality of a bootstrap distribution do we need to look for to be able
to construct confidence intervals using the standard error method?

---

## Section 8.4 

### `r emo::ji("computer")` R tutorial -- Currently experiencing issues, check back soon!

Now do Sections 3 and 4! <https://rstudio.csm.calpoly.edu/infer_CI_tutorial/>

---

## Section 8.5

### `r emo::ji("check")` Check-in: Interpreting a Confidence Interval

1. True or False -- We almost never know if our confidence interval captured the
true population parameter. 

2. What percentage of 99% confidence intervals do you expect to capture the 
true population parameter? 

<!-- use textbox -->

3. The word "confident" in a confidence interval interpretation corresponds to 
what aspect of the interval?

- the reliability of the procedure for constructing confidence intervals
- the accuracy of the original sample
- the precision of our bootstrap statistics

4. Which of the following are true?

- Lower confidence levels tend to produce narrower confidence intervals.  
- Lower confidence levels tend to produce wider confidence intervals.  
- Smaller sample sizes tend to produce wider confidence intervals. 
- Smaller sample sizes tend to produce narrower confidence intervals. 

---

## Section 8.7

### `r emo::ji("paper")` Group Concept Questions

**_Question 10_**

What is one similarity and one difference between using a bootstrap 
distribution to find a confidence interval versus using a t-distribution? 

<!-- Sim: -->
<!-- Both approximate the true sampling distribution -->

<!-- Diff: -->
<!-- Bootstraps resample to approximate the distribution, t-distributions use  -->
<!-- mathematical properties to approximate the distribution -->

---

# Think Out Loud Recording 

<!-- Something about sampling variability and sampling distributions -->

What is the difference between an accurate and a precise estimate? How do we
ensure that an estimate is accurate? How do we ensure that an estimate is
precise? 

