---
title: "Week One: Foundations of Statistics"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
library(flair)
library(emo)

library(unvotes)
library(lubridate)
library(DT)
```


Welcome!

In this coursework, you'll get a refresher on the foundational components of 
statistics and data, investigate how statistics is used in your major, and
think critically about the philosophy of statistical inference.

## Learning Outcomes 

By the end of this coursework you should be able to:

- describe observations, variables, and data matrices 
- explain the different types of variables a study can have 
- illustrate the difference between explanatory and response variables

- delineate the difference between a population and a sample
- compare and contrast different sampling methods
- outline what is necessary to make an experiment an "experiment" 
- characterize the differences between observational studies and experiments

- describe the difference between little "s" statistics and big "S" Statistics
- outline what "dichotomous thinking" is 
- illustrate how dichotomous thinking is used in Statistics
- explain issues with using the term "significant" in Statistics


## This Week's TODOs

`r emo::ji("play button")` Watch Video: 9 minutes

`r emo::ji("book")` Readings: 60-75 minutes 

`r emo::ji("computer")` Activities: 15 minutes 

`r emo::ji("check")` Check-ins: 2 (submit on Canvas)

`r emo::ji("scroll")` Group Discussion Homework: 6 Questions

---

# Data Refresher 

Let's start off by reading a refresher on some of the foundational concepts for 
data! 

`r emo::ji("book")` [**Required Reading:** _Introduction to Modern Statistics_ -- Hello Data](https://openintro-ims.netlify.app/data-hello.html)

`r emo::ji("book")` [**Required Reading:** _Introduction to Modern Statistics_ -- Study Design](https://openintro-ims.netlify.app/data-design.html)

`r emo::ji("book")` [**Required Reading:** statistics versus Statistics ](https://critical-inference.com/statistics-vs-statistics/)


`r emo::ji("check")` **Check-in 1:** 

1. What are the different types of variables that can appear in a 
dataset? (select all that apply)

- discrete numerical
- ordinal categorical
- nominal categorical 
- continuous numerical
- discrete categorical

2. Just because a variable has numeric values, does not mean it is a numeric
variable. How can you determine if a variable is numerical versus categorical? 

3. Which of the following statements are true about observational studies and
experiments? (select all that apply)

- Experiments randomly assign the explanatory variable
- Observational studies randomly assign the explanatory variable
- Observational studies can make causal statements about the relationship
between the explanatory and response variables
- Experiments can make causal statements about the relationship between the 
explanatory and response variables

4. What are different methods for sampling from a population? (select all that apply)

- simple random sampling 
- stratified random sampling 
- cluster sampling 
- multistage sampling
- convenience sampling

5. What does the term "stratification" in a *stratified random sample* mean? 

</br>

---

# Statistics in Your Field

`r emo::ji("play button")` [**Required Video:** How to Find a Peer Reviewed Journal Article (6 minutes)](https://www.youtube.com/watch?v=ui8ibJ0Pkhs)

Now let's investigate how statistics is used in your discipline. 

Find a peer-reviewed article from a journal in your field that performs a 
statistical analysis somewhere in the paper. You **do not** need to be familiar
with the analysis they do. That being said, if you want to find a paper that
uses a statistical method we'll cover in this class, look for a paper that uses
"regression" or "ANOVA." 

The Library is a great resource for this! The following website [https://csu-calpoly.primo.exlibrisgroup.com/discovery/jsearch?vid=01CALS_PSU:01CALS_PSU](https://csu-calpoly.primo.exlibrisgroup.com/discovery/jsearch?vid=01CALS_PSU:01CALS_PSU) has a list of journals by discipline that are available through the
library. 

`r emo::ji("check")` **Check-in 2:** Upload a PDF of your journal article to 
the Canvas assignment portal. 

</br>

---

# The Philosophy of Statistical Inference 

`r emo::ji("play button")` [**Required Video:** Philosophy of Statistics (3 minutes)](https://www.youtube.com/watch?v=K511BpyoTzU)

`r emo::ji("book")` [**Required Reading:** Our love of dichotomizing](https://critical-inference.com/our-love-of-dichotomizing/)

`r emo::ji("book")` [**Required Reading:** Science is Human](https://critical-inference.com/science-is-human/)

`r emo::ji("book")` [**Required Reading:** _Do We Really Need the S-word?_](https://www.americanscientist.org/article/do-we-really-need-the-s-word)

</br>

`r emo::ji("scroll")` **Group Homework Questions**

Throughout the week, your group will be asked to discuss the following 
questions:

- What is **S**tatistics?

- What are two similarities and two differences between Statistics and 
Mathematics? 

- What is the role of Statistics in the progression of science?

- How is dichotomous thinking used in Statistics? What are two 
benefits and drawbacks of dichotomous thinking?

- What does it mean for an analysis to be "objective"? Do you believe 
statistical analyses can be objective? Why or why not?

- What are 4-5 synonyms of "significant" that come to your mind?

- What are two issues that you see with using the term "significant" 
to denote "small" p-values? 

</br> 

The reporter in your team will be expected to type up your team's responses to 
these questions. The editor in your team is responsible for submitting the 
completed questions to the Canvas assignment portal. 
