---
title: "Data Visualization & Data Summary R Skills Review"
date: "January 19, 2021"
output: ioslides_presentation
fig_height: 4
fig_width: 4
---

<style type="text/css">
slides > slide:not(.nobackground):after {
  content: '';
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                       message = FALSE)

library(palmerpenguins)
library(tidyverse)
library(flair)
library(gridExtra)
library(ggridges)
```

## Data

Data were collected on 344 penguins living on three islands
(Torgersen, Biscoe, and Dream) in the Palmer Archipelago, Antarctica. In 
addition to which island each penguin lives on, the data contains information on
the species of the penguin (Adelie, Chinstrap, or Gentoo), its bill length, bill
depth, and flipper length (measured in millimeters), its body mass (measured in
grams), and the sex of the penguin (female or male).

## Data Wrangling

- 6 Verbs of Data Wrangling 
  * `filter()` -- filter out/in observations/rows
  * `select()` -- select variables/columns
  * `mutate()` -- modifies a dataset
  * `group_by()` -- group data by a categorical variable 
  * `summarize()` -- create data summaries of variables
  * `arrange()` -- arrange data in descending/ascending order

## `filter()`

## `select()` 

## `mutate()`

Adds new columns to a dataset **or** changes existing columns in a dataset. 

## `group_by()` 

## `summarize()` 

## `arrange()`

## Connecting Steps with the Pipe Operator 

```{r wrangle, include = FALSE}
penguins %>%
  filter(sex == "male", year == 2007) %>%
  group_by(species) %>%
  summarize(mean_bill_length = mean(bill_length_mm),
            sd_bill_length = sd(bill_length_mm),
            obs = n()) %>%
  arrange(desc(obs))
```

```{r, echo = FALSE}
decorate("wrangle") %>%
  flair("%>%")
```

```{r, echo = FALSE}
penguins %>%
  filter(sex == "male", year == 2007) %>%
  group_by(species) %>%
  summarize(mean_bill_length = mean(bill_length_mm),
            sd_bill_length = sd(bill_length_mm),
            obs = n()) %>%
  arrange(desc(obs))
```


## Histogram

```{r, message = FALSE, fig.height = 3, fig.width = 3}
penguins %>% 
  ggplot(aes(x = bill_length_mm)) + 
  geom_histogram()
```

## Dotplot

```{r, fig.height = 3, fig.width = 3}
penguins %>% 
  ggplot(aes(x = bill_length_mm)) + 
  geom_dotplot()
```

## Boxplot

```{r, fig.width = 3, fig.height = 5, echo = FALSE}
penguins %>% 
  ggplot(aes(x = bill_length_mm)) + 
  geom_boxplot()

```

## Density Plot 

```{r, fig.width = 4, fig.height = 4}
penguins %>% 
  ggplot(aes(x = bill_length_mm)) + 
  geom_density()
```

## Side-by-Side Boxplots

```{r, fig.width = 4, fig.height = 4, fig.align = "center"}
penguins %>% 
  ggplot(aes(y = bill_length_mm, x = species)) + 
  geom_boxplot()
```

## Side-by-Side Density Plots -- Ridge Plots

```{r, fig.width = 4, fig.height = 4, fig.align = 'center'}
penguins %>% 
  ggplot(aes(x = bill_length_mm, y = species)) + 
  geom_density_ridges()
```

## Scatterplots

```{r, fig.width = 4, fig.height = 4, fig.align = 'center'}
penguins %>% 
  ggplot(aes(y = bill_length_mm, x = bill_depth_mm)) + 
  geom_point()
```

## Colors in Scatterplots

```{r, fig.width = 6, fig.height = 4, fig.align = 'center'}
penguins %>% 
  ggplot(aes(y = bill_length_mm, x = bill_depth_mm, color = species)) + 
  geom_point()
```

## Colors in Ridge Plots

```{r, fig.width = 6, fig.height = 4, fig.align = 'center'}
penguins %>% 
  ggplot(aes(x = bill_length_mm, y = species, fill = species)) + 
  geom_density_ridges()
```

## Facets in Scatterplots 

```{r, fig.width = 8, fig.height = 4, fig.align = 'center'}
penguins %>% 
  ggplot(aes(y = bill_length_mm, x = bill_depth_mm, color = sex)) + 
  geom_point() + 
  facet_wrap(vars(species))
```

## Facets in Violin Plots

```{r, fig.width = 6, fig.height = 4, fig.align = 'center'}
penguins %>% 
  ggplot(aes(x = bill_length_mm, y = species, fill = sex)) + 
  geom_violin() + 
  facet_wrap(vars(island))
```
