---
title: "Simple Linear Regression -- R Skills Review"
date: "January 26, 2021"
output: ioslides_presentation
fig_height: 4
fig_width: 4
---

<style type="text/css">
slides > slide:not(.nobackground):after {
  content: '';
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      eval = FALSE, 
                      fig.align = "center")

library(openintro)
library(tidyverse)
library(flair)
library(gridExtra)
library(ggridges)
```

## Visualize & Summarize a Simple Linear Regression 

1. Visualize the relationship between x & y -- scatterplot 
2. Visualize the linear regression line -- `geom_smooth()`
3. Investigate unusual observations -- `filter()`

## Fit a Linear Regression Model

1. Calculate the correlation between x & y -- `summarize()`
2. Model the relationship between x & y -- `lm()`
3. Explore fitted values, residuals, etc. -- `augment()`
4. Make predictions for new data -- `predict()` 

## Data

The `ncbirths` dataset is a random sample of 1,000 cases taken from a larger 
dataset collected in North Carolina in 2004. 

Each case describes the birth of a single child born in North Carolina, along
with various characteristics of the child (e.g. birth weight, length of
gestation, etc.), the child’s mother (e.g. age, weight gained during pregnancy,
smoking habits, etc.) and the child’s father (e.g. age). 

##

```{r}
ncbirths

## relationship between weeks and weight 
ncbirths %>% 
  filter(weeks > 27) %>% 
  ggplot(aes(y = weight, x = weeks)) + 
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE)

## fit a linear regression
## explore the estimates from the regression 
## make a prediction from the regression 

birth_lm <- lm( weight ~ weeks + smoking, data = ncbirths)

library(broom)

new_births <- data.frame(weeks = c(42, 43, 44)) 

predict(birth_lm, newdata = new_births)



```





