---
title: "Randomization Tests for an ANOVA"
output: ioslides_presentation
---

<style type="text/css">
slides > slide:not(.nobackground):after {
  content: '';
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE)
library(tidyverse) 
library(palmerpenguins)
library(gridExtra)
library(openintro)
library(skimr)
library(infer)
library(okcupiddata)

classdata <- classdata %>% 
  mutate(exam = lecture)
```

## Goal of a Hypothesis Test for Many Means

An ANOVA seeks to assess the differences in the group means relative to the 
variability in the observations within each group. 

</br> 

Ideally, we'd like to see:  

- large differences in the means of the groups 
- small amounts of variability within each group


## Hypotheses 

ANOVA uses a single hypothesis test to check whether the means across many
groups are equal:

$H_0$: The mean outcome is the same across all groups.  

_OR_ 

$H_0$: $\mu_1 = \mu_2 = \cdots = \mu_k,$  where $\mu_i$ represents the
mean of the response for observations in category $i$.

</br>

$H_A$: At least one mean ($\mu_i$) is different.  

## Model Conditions 

Before performing an ANOVA, there are three conditions we must check: 

- the observations are independent within and across groups  

- the responses within each group are nearly normal 

- the variability across the groups is similar

## Calculating an Observed Statistic

We need a statistic that captures:  

1. how different the group of means are from each other 
2. how variable the observations of the groups are  

</br> 

The $F$-statistic does just this!   

It is the ratio of the variability *between* the groups to the variability
*within* the groups. 

$$ F = \frac{MSG}{MSE} $$


## Null Hypothesis

An ANOVA focuses on answering one question: is the variability in the sample
means so large that it seems unlikely to be from chance alone?

</br> 

If the null hypothesis is true, any differences among the sample means are only
due to chance. This would mean that the $MSG$ and $MSE$ should be about equal.
  
  - This means, if the null hypothesis is true, we'd expect an $F$-statistic of 
  about 1. 
  

## Exam Scores

Suppose now that the teacher had had such an extremely large class that three
different exams were given: A, B, and C. We’d like to investigate whether or not
the difficultly of the exams is the same across the three exams. 

So, our test would be: 

$H_0$: $\mu_A = \mu_B = \mu_C$; The inherent average difficulty is the same
across the three exams.  

$H_A$: at least one $\mu_i$ (i = A, B, C) is different; At least one of the
exams is inherently more (or less) difficult than the others.


## Exploratory Visualization

```{r, fig.height=3.5, fig.width=4, fig.align='center'}
classdata %>% 
  ggplot(aes(x = exam, y = m1)) + 
  geom_violin(aes(fill = exam)) + 
  geom_jitter(width = 0.2) + 
  theme(legend.position = "none")
```

## Randomization Test

- If the null hypothesis is true, then the score on each exam should represent
the true student ability on that material. 

- It shouldn't matter whether a student took exam A or exam B or exam C. 

- By reallocating which student got which exam, we are able to understand how
the difference in average exam scores changes due only to natural variability. 

## Using infer 

We have the same steps as before, the only modifications are to the statistic
calculated for each permutation. 

```{r}
null_dist <- classdata %>% 
  specify(m1 ~ exam) %>% 
  hypothesise(null = "independence") %>% 
  generate(reps = 2000, type = "permute") %>% 
  calculate(stat = "F")
```

## Visualizing the Null Distribution 

<div class="columns-2">

```{r, eval = FALSE}
obs_F <- classdata %>% 
  specify(m1 ~ exam) %>% 
  calculate(stat = "F")

null_dist %>% 
  visualise() + 
  geom_vline(xintercept = obs_F)
```

</br> 
</br> 
</br> 
</br> 
</br> 
</br> 

```{r, echo = FALSE, fig.width=4, fig.height=4}
obs_F <- classdata %>% 
  specify(m1 ~ exam) %>% 
  calculate(stat = "F") %>% 
  pull()

null_dist %>% 
  visualise() + 
  shade_p_value(obs_stat = obs_F, 
                direction = "greater")
```

## Calculating the p-value

```{r}
null_dist %>% 
  get_p_value(obs_stat = obs_F, direction = "greater")
```

</br> 

- What would you conclude? 

- Which exam seemed different from the others? 


