---
title: Inference for to Linear Regression
author: YOUR NAME HERE!
date: "Due: Sunday, November 21 by 11:59pm"
output: html_document
---

```{r global_options, include=FALSE}
## These are global specifications for how the RMarkdown file should knit. 
## Do not change these or delete these!
knitr::opts_chunk$set(eval = TRUE,
                      warning = FALSE,
                      fig.align = "center", 
                      message = FALSE, 
                      fig.width = 4, 
                      fig.height = 4)
```

The Human Freedom Index is a report that attempts to summarize the idea of "freedom" through a bunch of different variables for many countries around the globe.
It serves as a rough objective measure for the relationships between the different types of freedom - whether it's political, religious, economical or personal freedom - and other social and economic circumstances.
The Human Freedom Index is an annually co-published report by the Cato Institute, the Fraser Institute, and the Liberales Institut at the Friedrich Naumann Foundation for Freedom.

In this lab, you'll be analyzing data from Human Freedom Index reports from 2008-2016.
Your aim will be to summarize a few of the relationships within the data both graphically and numerically in order to discuss how different variables can help tell a story about freedom.

# Getting Started

## Load Packages

In this lab, you will explore and visualize the data using the **tidyverse** suite of packages. 

You will also use the **moderndive** package to get nice looking regression tables. 

The data can be found in the **openintro** package.

Let's load the packages!

```{r load-packages, message=FALSE}
library(tidyverse)
library(moderndive)
library(openintro)
```

# The data

The data we're working with is in the openintro package and it's called `hfi`, short for Human Freedom Index.

__1.  The dataset has **a lot** of variables, but we are only interested in four: hf_score, pf_score, year, and region. *Select* these variables from the hfi dataframe, and assign the result to a dataframe named hfi_small.__ 

```{r}
## Fill in the code necessary for question 1 

```


We can use the `drop_na()` function to get an idea of how many observations have missing data. The function looks row by row through the data to decide what rows should be kept. If there is an `NA` value in a row, that row is discarded. The result is a dataframe where there are no missing values for any of the variables. 

__2. Use the drop_na() function to create another new dataframe named hfi_small_clean that **does not** contain any missing values.__ 

```{r}
hfi_small_clean <- drop_na(____)

```

# Visualizing the Relationship Between Economic Freedom and Personal Freedom 

__3. (a) Plot the relationship between `ef_score` (x) and `pf_score` (y), including a linear regression line.__
__3. (b) Does the relationship look linear? How strong is the relationship?__
Hint: This is a great instance to use `geom_jitter()`!  
<!-- Your response goes here! Do not remove this line! -->


```{r}
## Code for question 3

```

__4. Now, add a facet for each year in the dataset. Does it appear that this relationship changes over time?__

<!-- Change your code from Question 4. Only your response goes here! -->


# Modeling the Relationship Between Human Freedom and Personal Freedom 

We will investigate the relationship between Personal Freedom and Economic 
Freedom, determining if there is evidence that this relationship changes over
time. 

Our statistical method will be driven by the conditions for our linear
regression. 

## Checking Conditions 

You've already discussed the *linear relationship* condition, so let's move on 
to the independence condition. 

__5. Given that we're accounting for year, do you believe it is reasonable to assume these observations are independent? Why or why not?__
<!-- Your response goes below this line! -->


To perform a residual analysis, we first need to fit the multiple linear 
regression. We are interested in determining if the relationship between 
pf_score and ef_score is different for different years. This should sound similar
to the interaction model we used before! 

__6. Fill in the code below to fit a multiple linear regression, which allows for an interaction between year and ef_score.__

```{r}
pf_lm <- lm(hf_score ~ ____, data = hfi)
```

Next, you'll use the code below to diagnose the *normality* and *equal variance* conditions. 

```{r}
get_regression_points(pf_lm) %>% 
  ggplot(mapping = aes(x = residual)) + 
  geom_histogram(binwidth = 0.15)
```

__7. Based on the plot above, do you believe the normality condition is violated? Why or why not?__
<!-- Your response goes here! Do not remove this line! -->



```{r}
get_regression_points(pf_lm) %>% 
  ggplot(mapping = aes(y = residual, x = hf_score_hat)) + 
  geom_jitter() + 
  geom_hline(yintercept = 0, color = "red", size = 1)

```

__8. Based on the plot above, do you believe the equal variance condition is violated? Why or why not?__
<!-- Your response goes here! Do not remove this line! -->



## Hypothesis Test for the Slope 

First, we will carry out a hypothesis test to investigate if the relationship
between hf_score and pf_score differs by year. This entails that we are first 
testing to see if we need the interaction term in our multiple linear 
regression. These hypotheses are a bit different than before, so let's walk
through them. 

To test if an interaction term is needed, we are testing if the slope on
pf_score should change based on the year. For a quantitative variable, this is
either a "yes" or "no" answer. The hypotheses look like the following:

H_0: the true relationship between x and y *does not* differ by z

H_A: the true relationship between x and y *does* differ by z


__9. Using the template above, write out the null and alternative hypotheses we are testing, in the context of these data.__   
<!-- Your response goes here! Do not remove this line! -->




__10. Report the p-value for this hypothesis test. What would you conclude for this test?__   
Hint 1: Deciding to reject or not **is not** a conclusion, you need to specify which hypothesis seems more plausible **and** what that means in the context of these data!   
Hint 2: Stating you reject the null hypothesis requires that you specify the significance threshold that was used.
Hint 3: There should be no use of the term "significant" in your conclusion. 
<!-- Your response goes here! Do not remove this line! -->



```{r}
## Code for finding the p-value goes below this line! 


```


__11. How would you interpret this interaction term?__
__Give this your best shot! I will be grading based on the ideas you present.__
__See if you can make connections between this coefficient and what you saw with a categorical explanatory variable.__
<!-- Your response goes here! Do not remove this line! -->



