---
title: "Two-Way ANOVA -- A Brief Introduction"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", 
                      fig.height = 3, 
                      fig.width = 5,
                      echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE)

library(tidyverse) 
library(palmerpenguins)
library(gridExtra)
library(ggridges)
library(openintro)
library(infer)
library(learnr)
library(broom)


penguins <- penguins %>% 
  mutate(year_cat = as.factor(year))
```

## Two-way ANOVA

A two-way ANOVA extends the one-way ANOVA to situations with two categorical
explanatory variables. This new methods allows researchers to simultaneously
study two variables that might explain variability in the responses and explore
whether the impacts of one explanatory variable change depending on the level of
the other explanatory variable.

In a clinical trials context, it is well known that certain factors can change
the performance of certain drugs. For example, different dosages of a drug might
have different benefits or side-effects on men, versus women or children or even
for different age groups in adults. When the impact of one factor on the
response changes depending on the level of another factor, we say that the two
explanatory variables __interact__. 

It is also possible for both factors to be related to differences in the mean
responses and not interact. For example, suppose there are differences in how 
younger and older subjects respond to a drug, and there are differences in how 
all individuals respond to different dosages of a drug, but the effect of
increasing the dosage is the same for both young and old subjects. This is an
example of what is called an __additive__ type of model. 

In general, the world is more complicated than the single factor models we've 
considered, especially in observational studies, so these models allow us to
start to handle more realistic situations.


## Visualizing a Two-Way ANOVA

The visualizations we created for a one-way ANOVA are still relevant here, but 
we need to figure out how to add a second categorical explanatory variable to 
our plots. 

Similar to the multivariate plots we've talked about previously, there are 
two main ways to add a second categorical variable to our plots:  

1. colors  
2. facets 

We'll explore both below! 

## 

Below is a plot of the relationship between flipper length and species for 
penguins in the Palmer Archipelago. 

The year the data were collected was modified to a categorical variable, named  
`year_cat`. Change to code below to `fill` the boxplots and density ridges 
with the `year_cat` variable. 

```{r colors, exercise = TRUE}
penguins %>% 
  ggplot(aes(y = species, x = flipper_length_mm, fill = ___)) + 
  geom_boxplot()

penguins %>% 
  ggplot(aes(y = species, x = flipper_length_mm, fill = ___)) + 
  geom_density_ridges()

```

```{r colors-solution, include = FALSE}
ggplot(aes(y = species, x = flipper_length_mm, fill = year_cat)) 
```

##

Now, take the same plot and instead of coloring by `year_cat` use facets to 
separate the different sampling years. 

```{r facets, exercise = TRUE}
penguins %>% 
  ggplot(aes(y = species, x = flipper_length_mm)) + 
  geom_boxplot() + 
  ___

penguins %>% 
  ggplot(aes(y = species, x = flipper_length_mm)) + 
  geom_density_ridges() + 
  ___
```

<div id="facets-hint">
**Hint:** Add `vars(year_cat)` to `facet_wrap()` to create year facets. 
</div>

```{r facets-solution, include = FALSE}
penguins %>% 
  ggplot(aes(y = species, x = flipper_length_mm)) + 
  geom_boxplot() + 
  facet_wrap(vars(year_cat))

penguins %>% 
  ggplot(aes(y = species, x = flipper_length_mm)) + 
  geom_density_ridges() + 
  facet_wrap(vars(year_cat))
```

## Additive versus Interactive Models 

As was mentioned in the Introduction, there are two different types of two-way 
ANOVA models. Similar to a multiple linear regression, the two explanatory 
variables could have their own impact on the response (similar to a parallel 
slopes regression model). Or, the relationship between one explanatory variable
and the response could differ based on another explanatory variable (similar to 
a different slopes regression model). 

These two types of models are called an __additive__ two-way ANOVA model or an 
__interaction__ two-way ANOVA model. Similar to how we decided which model to 
choose in a multiple linear regression, we will use visualizations to guide us. 

## 

When deciding if an interaction model is a good fit for the data, we look to see 
if the relationship between one categorical variable and the response differs 
based on the level of the other response variable. 

Here, we look at the relationship between `year_cat` and `flipper_length_mm` and
see if it differs based on the species of the penguins. I think of looking at 
the shape of the relationships inside of the outlined box. I draw a line through 
the center of the Adelie boxplots. I then look to see if it seems like the slope
of the line through the Adelie penguins looks different from the slope of the
lines drawn through the Chinstrap and Gentoo penguins. 

```{r}
penguins %>% 
  ggplot(aes(x = species, y = flipper_length_mm)) + 
  geom_boxplot(aes(fill = year_cat)) +
  geom_segment(aes(x=0.5, xend=1.5, y=170, yend=170), linetype = 2, size = 1) + ## BOTTOM
  geom_segment(aes(x=0.5, xend=1.5, y=210, yend=210), linetype = 2, size = 1) + ## TOP
  geom_segment(aes(x=0.5, xend=0.5, y=170, yend=210), linetype = 2, size = 1) +
  geom_segment(aes(x=1.5, xend=1.5, y=170, yend=210), linetype = 2, size = 1) 

```

Does it appear that the relationship between year and flipper length differs based 
on the species of the penguin? I didn't think so, since it seems like the slope 
of the line through the three years is about the same for all three species. 
However, we can use an F-test to check and see if my intuition is right. 

## Conditions of a Two-Way ANOVA Model 

The two-way ANOVA model has the same conditions as its one-way counterpart, 
however, we now have extra groups to think about. 

__Independence__  
For both categorical variables:  

- observations across groups need to be independent

- observations within each group need to be independent 
 
__Equal Variance__:  the variability of each group is similar to the others.   

*Note:* this is an 
assumption about the groups at the intersection of each categorical variable 
(e.g. Gentoo penguins captured in 2007). This assumption is fairly robust, but 
large differences in variability will cause issues. 

__Normal Distribution__: the responses of each group need to be approximately Normal.   

*Note:* this is an 
assumption about the groups at the intersection of each categorical variable 
(e.g. Gentoo penguins captured in 2007). This assumption is 
also fairly robust, but influential outliers and the sample 
sizes of the groups should be noted when assessing. 

## 

Your turn! 

```{r quiz_1}
question("Which of the following is important to consider when assessing independence?",
  correct = "Correct! We need to look across groups and within groups. ",
  allow_retry = TRUE,
  answer("If the same penguins could be sampled multiple years.", correct = TRUE),
  answer("If the same penguins could be sampled at multiple islands.", message = "No, island is not a variable we are interested in!"),
  answer("If penguins in each species could share similar information.", correct = TRUE),
  answer("If penguin species share similar characteristics", message = "No, this is not a consideration when assessing independence.")
)
```

```{r quiz_2}
question("Based on the plot below, is the condition of equal variablility violated?",
  correct = "Correct! The length of the violins looks similar across the different groups.",
  allow_retry = TRUE,
  answer("No.", correct = TRUE),
  answer("Yes.", message = "Give the plot another look. Does it appear that the length of the violins of the groups are substantiall different?"),
  answer("It's difficult to tell. ", message = "Does it appear that the width of the density plots of the groupsare substantially different?")
)
```

```{r}
penguins %>% 
  ggplot(aes(y = species, x = flipper_length_mm)) + 
  geom_density_ridges() + 
  facet_wrap(vars(year_cat))

```

```{r quiz_3}
question("Based on the same plot above, is the condition of normally distributed responses violated?",
  correct = "Correct! The distributions all look 'normal enough' and there are no large outliers in any of the groups.",
  allow_retry = TRUE,
  answer("Yes.", message = "Give the plot another look. This condition is fairly robust to slight differences in shape, but we do need to worry about outliers. Are there large outliers present in any of the groups?"),
  answer("No.", correct = TRUE),
  answer("It's difficult to tell.", message = "Are there large outliers present in any of the groups?")
)
```


## Fitting a Two-Way ANOVA Model 

The `R` code to fit a two-way ANOVA model is very similar to a one-way ANOVA model. 
We use the `aov()` (analysis of variance) function, but now we will have two 
categorical explanatory variables. 

First, we'll fit an interaction two-way ANOVA model to see if the relationship 
between year and flipper length differs based on species of penguins.   

- Similar to different slopes model, `species:year_cat` is the *interaction* between 
these two variables. 

```{r, echo = TRUE}
aov(flipper_length_mm ~ species + year_cat + species:year_cat, 
    data = penguins) %>% 
  tidy()
```

</br> 

We notice that the interaction line (`species:year_cat`) has a small F-statistic
(0.551) and a large p-value (0.698). This would lead for us to conclude that 
there is not an interaction between these two variables, or that the relationship 
between year and flipper length does not differ based on penguin species. 

## 

An additive model seems like a better choice. Modify the code from the
interaction model to fit an additive model instead. 


```{r additive, exercise = TRUE}
aov(flipper_length_mm ~ species + year_cat + species:year_cat, data = penguins) %>% 
  tidy()
```

<div id="additive-hint">
**Hint:** Remove the interaction term to fit an additive model! 
</div>


```{r additive-solution, include = FALSE}
aov(flipper_length_mm ~ species + year_cat, data = penguins) %>% 
  tidy()
```

</br> 

Based on the ANOVA table above, what would you conclude for the relationship 
between flipper length and species and the relationship between flipper length 
and sampling year? 

## Model Conclusions 

In this context, the ANOVA table allows for use to test two hypotheses: 

1. whether the mean flipper lengths for every year are equal 
2. whether the mean flipper lengths for every species are equal 

Again, similar to a multiple regression, the interpretation of these tests is 
*conditional* on the other variable in the model.

## 

Based on the ANOVA table below, we would conclude that, after accounting for the 
species of penguin, at least one year of capture has a different mean flipper
length. 

```{r}
aov(flipper_length_mm ~ species + year_cat, data = penguins) %>% 
  tidy()
```

```{r quiz_4}
question("Based on the ANOVA table above, what would you conclude for the second hypothesis test?",
  correct = "Correct!",
  allow_retry = TRUE,
  answer("Fail to reject H0, the mean flipper lengths of the years are the same, after accouting for uear of capture.", message = "Careful! Look at the p-value for the `species` line. Is is small? What does that say about whether it seems reasonable to say that the mean flipper lengths are the same for every species?"),
  
  answer("There is strong evidence that at least one species has a different mean flipper length, after accouting for year of capture.", correct = TRUE), 
  answer("There is strong evidence that each species has a different mean flipper length, after accouting for year of capture.", message = "Careful! What is the null hypothesis for an ANOVA? What is the opposite of saying 'everything is equal'?"), 
answer("There is strong evidence that at least one species has a different mean flipper length.", message = "Careful! What is this hypothesis test conditional on?") 
)
```
