---
title: "Introduction to Visualizing and Summarizing Data"
author: "Joshua Chang"
date: "Due: January 24, 2021 at 11:59pm"
output: html_document
---

<div style="color:blue">

Great job! You have revisions on problems: 

My comments for your responses to the exercises can be found in blue text, and my comments/suggestions for your R code can be found in the code chunks behind two hashtags (##)

Your homework assignment receives a score that reflects the average points-per-question, where each question is worth 4 points. This is assigned based on the grading system outlined in the syllabus. 

Problems marked with an “R” must be redone and returned within two days of receiving the assignment back. You will receive a score of 0 on each problem requiring a revision until a revision is submitted. Late revisions receive a zero, unless you contact me. Work successfully redone (on time) will receive a maximum grade of 3. It is possible to have a redo on a redo.

</div>


```{r, include = FALSE}
knitr::opts_chunk$set(echol = TRUE, 
                      results = "hold", 
                      message = FALSE, 
                      warning = FALSE)
```

## Getting started

### Load packages

In this lab, we will explore and visualize the data using the **tidyverse** suite of packages. The data can be found in the companion package for OpenIntro labs, **openintro**.

Let's load the following packages:

-   The **tidyverse** "umbrella" package which houses a suite of many different `R` packages: for data wrangling and data visualization
-   The **openintro** `R` package: for data and custom functions with the OpenIntro resources

```{r, include = FALSE}
library(tidyverse)
library(openintro)
## This is a comment -- R will not run this line as code, because it has ## before it
```

### The data

The [Bureau of Transportation Statistics](http://www.rita.dot.gov/bts/about/) (BTS) is a statistical agency that is a part of the Research and Innovative Technology Administration (RITA).
As its name implies, BTS collects and makes transportation data available, such as the flights data we will be working with in this lab.

First, we'll view the `nycflights` data frame.
Type the following in your console to load the data:

```{r}
data(nycflights)
```

The data set `nycflights` that shows up in your workspace is a *data matrix*, with each row representing an *observation* and each column representing a *variable*.
R calls this data format a **data frame**, which is a term that will be used throughout the labs.
For this data set, each *observation* is a single flight.

To view the names of the variables, you can use the `names()` function: 

```{r}
names(nycflights)
```

This returns the names of the variables in this data frame.
The **codebook** (description of the variables) can be accessed by pulling up the help file by typing a `?` before the name of the dataset:

```{r, eval = FALSE}
?nycflights
```

One of the variables refers to the carrier (i.e. airline) of the flight, which is coded according to the following system.

-   `carrier`: Two letter carrier abbreviation.

    -   `9E`: Endeavor Air Inc.
    -   `AA`: American Airlines Inc.
    -   `AS`: Alaska Airlines Inc.
    -   `B6`: JetBlue Airways
    -   `DL`: Delta Air Lines Inc.
    -   `EV`: ExpressJet Airlines Inc.
    -   `F9`: Frontier Airlines Inc.
    -   `FL`: AirTran Airways Corporation
    -   `HA`: Hawaiian Airlines Inc.
    -   `MQ`: Envoy Air
    -   `OO`: SkyWest Airlines Inc.
    -   `UA`: United Air Lines Inc.
    -   `US`: US Airways Inc.
    -   `VX`: Virgin America
    -   `WN`: Southwest Airlines Co.
    -   `YV`: Mesa Airlines Inc.

Remember that you can use `glimpse` to take a quick peek at your data to understand its contents better.

```{r}
glimpse(nycflights)
```

1. (a) How large is the `nycflights` dataset? (i.e. How many rows and columns does it have?)
<!-- Your response goes below this line! -->

The nycflights dataset has 32,735 rows and 16 columns, indicating that there are 32,735 observations with 16 variables for each observation.

<div style="color:blue">
Correct! Great specific description of what this means!  
</div>

1. (b) When were these data collected? 
<!-- Your response goes below this line! -->

These data were collected from a random sample of flights that departed NYC in 2013.

<div style="color:blue">
Correct! Nice use of the help file, now you know these are a random sample of flights!  
</div>


1. (c) Are there categorical variables in the dataset? If so, what are their names and data types? 
<!-- Your response goes below this line! -->

Yes, there are multiple categorical variables in the dataset: year, month, day, carrier, tailnum, flight, origin, and dest. 

Year, month, and day are ordinal categorical variables that depict the date of departure. Flight and tailnum are also ordinal categorical variables that depict details of the flight number and plane tail number, respectively. However, carrier, origin, and dest are nominal (not ordinal) categorical variables that depict flight information that cannot be inherently ordered.

<div style="color:blue">
Nice work! Keep in mind that if these variables were labeled a `fct` that would *also* be a categorical variable!  
</div>

<div style="color:blue">
Overall: 4
</div>


### Departure Delays

Let's start by examining the distribution of departure delays of all flights with a histogram.

```{r}
ggplot(data = nycflights, aes(x = dep_delay)) +
  geom_histogram()
```

This function says to plot the `dep_delay` variable from the `nycflights` data frame on the x-axis.
It also defines a `geom` (short for geometric object), which describes the type of plot you will produce.

Histograms are generally a very good way to see the shape of a single distribution of numerical data, but that shape can change depending on how the data is split between the different bins.
You can easily define the binwidth you want to use, by specifying the `binwidth` argument to `geom_histogram()`, like so:

```{r}
ggplot(data = nycflights, aes(x = dep_delay)) +
  geom_histogram(binwidth = 15)

ggplot(data = nycflights, aes(x = dep_delay)) +
  geom_histogram(binwidth = 150)
```

2.  Look carefully at these three histograms. How do they compare? Are features revealed in one that are obscured in another?
<!-- Your answer goes below this line -- do not remove this!  -->

Yes, increasing the width of the bins reduces the number of columns represented on the histogram, as indicated in the third histogram. This differs from the second and first histogram, which have more columns/bins as a result of smaller bin widths. To which, we can understand how bins that are too wide can obscure important details about the distribution of data. On the other hand, bins that are too narrow can reveal more data than needed that may also obscure data distribution. Having a correct bin width is necessary to show accurate data distributions, without producing too much noise or having too few columns that create obscurity.

<div style="color:blue">
Nice work! The smaller binwidth makes certain aspects of plot more visible. But you are also correct that very small binwidths give too much detail and are difficult to interpret.    
</div>

<div style="color:blue">
Overall: 4
</div>

3. A histogram is only one way to visualize the distribution of a numerical variable. 
(a) What are other visualizations that we could have produced? 
(b) What are the benefits and weaknesses of a histogram compared to these other options? 
<!-- Your response goes below this line! -->

As an alternative to histograms, we could have visualized the distribution of the numerical value by using a dotplot, density plot, or box plot. 

A histogram has multiple benefits over the other options because it is versatile in visualizing datasets. In comparison to a dotplot, histograms are easier to read for datasets of large samples by binning observation values into visual bars. In addition, we are visualize data densities in datasets as higher bars indicate more common occurrences of values. This is advantageous in trying to visualize and understand the overall shape of the data distribution. To which, we can even identify modes through histograms to better understand our data and compare the overall shape with normal curves. 

However, histograms have a weakness in that they struggle to compare more than two data sets and cannot read exact values as the data is grouped into categories. In addition, histograms may also present misleading data if bins are too large or too small, thereby exaggerating or minimizing the information within each bin.

<div style="color:blue">
Great description of the benefits and weaknesses of histograms, but how do these relate to the other plotting options? 
</div>

<div style="color:blue">
Overall: 3
</div>


## LAX Destinations 

If you want to visualize only on delays of flights headed to Los Angeles, you need to first `filter()` the data for flights with that destination (`dest == "LAX"`) and then make a histogram of the departure delays of only those flights.

```{r}
lax_flights <- nycflights %>%
  filter(dest == "LAX")

ggplot(data = lax_flights, aes(x = dep_delay)) +
  geom_histogram()
```

Let's decipher these two commands (OK, so it might look like four lines, but the first two physical lines of code are actually part of the same command. It's common to add a break to a new line after `%>%` to help readability).

-   Command 1: Take the `nycflights` data frame, `filter()` for flights headed to LAX, and save the result as a new data frame called `lax_flights`.

    -   `==` means "if it's equal to".
    -   `LAX` is in quotation marks since it is a character string.

-   Command 2: Basically the same `ggplot()` call from earlier for making a histogram, except that it uses the smaller data frame for flights headed to LAX instead of all flights.

**Logical operators:** Filtering for certain observations (e.g. flights from a particular airport) is often of interest in data frames where we might want to examine observations with certain characteristics separately from the rest of the data.
To do so, you can use the `filter()` function and a series of **logical operators**.
The most commonly used logical operators for data analysis are as follows:

-   `==` means "equal to"
-   `!=` means "not equal to"
-   `>` or `<` means "greater than" or "less than"
-   `>=` or `<=` means "greater than or equal to" or "less than or equal to"


4. You learned in the tutorial that this process of creating a new dataset based on a filter is not necessarily the best plan. An alternative plan is to use the pipe operator (`%>%`) to create a data pipeline, carrying the filtered data into the plot. 
Change the code below, so it uses the pipe operator instead of saving a new dataset. 

```{r}
## For this exercise, you need to change the code below! 

nycflights %>%
  filter(dest == "LAX") %>%
  ggplot(aes(x = dep_delay)) +
  geom_histogram()

```

<div style="color:blue">
Great use of the pipe operator!
</div>

<div style="color:blue">
Overall: 4
</div>


### Multiple Data Filters 

You can also filter based on multiple criteria.
Suppose you are interested in flights headed to San Francisco (SFO) in February:

```{r}
sfo_feb_flights <- nycflights %>%
  filter(dest == "SFO", month == 2)
```

Note that you can separate the conditions using commas if you want flights that are both headed to SFO **and** in February.
If you are interested in either flights headed to SFO **or** in February, you can use the `|` instead of the comma.


5.  Find the number of flights in this dataset that were headed to SFO in February. 
**Hint:** The `filter()` and `count()` functions will be helpful!  

```{r}
## Code for exercise 1 here! 
nycflights %>%
  filter(dest == "SFO", month == 2) %>%
  count()

```

<div style="color:blue">
Nice work! 
</div>

<div style="color:blue">
Overall: 4
</div>
The dataset contains 68 flights that were headed to SFO in February.

## Data Summaries 

You can also obtain numerical summaries for the flights headed to LAX, using the `summarise()` function:

```{r}
lax_flights %>%
  summarise(mean_dd   = mean(dep_delay), 
            median_dd = median(dep_delay), 
            n         = n())
```

Note that in the `summarise()` function you created a list of three different numerical summaries that you were interested in.
The names of these elements are user defined, like `mean_dd`, `median_dd`, `n`, and you can customize these names as you like (just don't use spaces in your names).
Calculating these summary statistics also requires that you know the summary functions you would like to use. 

**Summary statistics:** Some useful function calls for summary statistics for a single numerical variable are as follows:

-   `mean()`: calculates the average
-   `median()`: calculates the median
-   `sd()`: calculates the standard deviation
-   `var()`: calculates the variances
-   `IQR()`: calculates the inner quartile range (Q3 - Q1)
-   `min()`: finds the minimum 
-   `max()`: finds the maximum
-   `n()`: reports the sample size

Note that each of these functions takes a single vector as an argument and returns a single value.

## Data Summaries & Data Visualizations 

6.  Describe the distribution of the **arrival** delays of the LAX flights using a histogram and appropriate summary statistics.
**Hint:** The summary statistics you use should depend on the shape of the distribution.
<!-- Your response to this problem goes below this line! -->

<div style="color:blue">
Great visualization and data summaries! Are the summaries you found appropriate for this distribution?  
Where is your description? 
</div>

<div style="color:blue">
Overall: Redo
</div>


```{r}
## Code for exercise 2 goes here! 
ggplot(data = lax_flights, aes(x = arr_delay)) +
    geom_histogram()

lax_flights %>%
  summarise(mean_ad   = mean(arr_delay), 
            median_ad = median(arr_delay), 
            sd_ad         = sd(arr_delay))

```

## Grouped Summaries 

Another useful technique is quickly calculating summary statistics for various groups in your data frame.
For example, we can modify the above command using the `group_by()` function to get the same summary stats for each origin airport:

```{r}
sfo_feb_flights %>%
  group_by(origin) %>%
  summarise(median_dd = median(dep_delay), 
            iqr_dd = IQR(dep_delay), 
            n_flights = n())
```

Here, we first grouped the data by `origin` and then calculated the summary statistics.

## Summaries vs. Visualizations 

Which month would you expect to have the highest average delay departing from an NYC airport?

Let's think about how you could answer this question. 
One option is to summarize the data and inspect the output. 
Another option is to plot the delays and inspect the plots. 
Let's try both! 

7. Find the average departure delays for each month, and arrange these average delays in descending order. 
**Hint:** the `desc()` function is helpful for arranging numbers in descending order! 

```{r}
## Code for exercise 3 goes here!
nycflights %>%
  group_by(month) %>%
  summarise(mean_dd = mean(dep_delay)) %>%
  arrange(desc(mean_dd))



```

<div style="color:blue">
Nice work combining all of these pieces together! Woohoo! Data rodeo! 
</div>

<div style="color:blue">
Overall: 4
</div>


8. Now, plot the distribution of departure delays, faceted by the month of departure. 
Choose the type of plot you believe is appropriate for visualizing the distribution of departure delays. 

```{r}
## Code for exercise 4 goes here! 
nycflights %>%
  ggplot(aes(x = dep_delay)) +
  geom_histogram() +
  facet_wrap(vars(month))

```

<div style="color:blue">
Nice work faceting by month! 
</div>

<div style="color:blue">
Overall: 4
</div>


9. What information can you obtain from the visualization that you could not from the data summaries?   
<!-- Your response to this question goes below this line!  -->

Though the data summaries would indicate that July would have the highest average of flight delays, it does not indicate the distribution of flight delays that occur in each month. This is important as understanding the distribution allows for more practical uses of the data. In this case, the visualization provides us with the knowledge that some months may be more likely to have flight delays than others as well as how long those delays will be. I believe it is also important to note that because of outliers, the visualization is a little more difficult to read. However, if extreme outliers were taken out of the dataset prior to visualization, then we would have much better views of each month's distribution. To which, we can also understand that the mean may also not be the best measure of center as it can be sensitive to extreme values. This is something that we may not be able to see without adding in the median value of each month's dep_delay variable or without using a visualization.
