---
title: "Introduction to Visualizing and Summarizing Data"
author: "Your Name Here!"
date: "Due: January 24, 2021 at 11:59pm"
output: html_document
---

<div style="color:blue">

Great job! You have revisions on problems: 

My comments for your responses to the exercises can be found in blue text, and my comments/suggestions for your R code can be found in the code chunks behind two hashtags (##)

Your homework assignment receives a score that reflects the average points-per-question, where each question is worth 4 points. This is assigned based on the grading system outlined in the syllabus. 

Problems marked with an “R” must be redone and returned within two days of receiving the assignment back. You will receive a score of 0 on each problem requiring a revision until a revision is submitted. Late revisions receive a zero, unless you contact me. Work successfully redone (on time) will receive a maximum grade of 3. It is possible to have a redo on a redo.

</div>


```{r, include = FALSE}
knitr::opts_chunk$set(echol = TRUE, 
                      results = "hold", 
                      message = FALSE, 
                      warning = FALSE)
```

## Getting started

### Load packages

In this lab, we will explore and visualize the data using the **tidyverse** suite of packages. The data can be found in the companion package for OpenIntro labs, **openintro**.

Let's load the following packages:

-   The **tidyverse** "umbrella" package which houses a suite of many different `R` packages: for data wrangling and data visualization
-   The **openintro** `R` package: for data and custom functions with the OpenIntro resources

```{r, include = FALSE}
library(tidyverse)
library(openintro)
## This is a comment -- R will not run this line as code, because it has ## before it
```

### The data

The [Bureau of Transportation Statistics](http://www.rita.dot.gov/bts/about/) (BTS) is a statistical agency that is a part of the Research and Innovative Technology Administration (RITA).
As its name implies, BTS collects and makes transportation data available, such as the flights data we will be working with in this lab.

First, we'll view the `nycflights` data frame.
Type the following in your console to load the data:

```{r}
data(nycflights)
```

The data set `nycflights` that shows up in your workspace is a *data matrix*, with each row representing an *observation* and each column representing a *variable*.
R calls this data format a **data frame**, which is a term that will be used throughout the labs.
For this data set, each *observation* is a single flight.

To view the names of the variables, you can use the `names()` function: 

```{r}
names(nycflights)
```

This returns the names of the variables in this data frame.
The **codebook** (description of the variables) can be accessed by pulling up the help file by typing a `?` before the name of the dataset:

```{r, eval = FALSE}
?nycflights
```

One of the variables refers to the carrier (i.e. airline) of the flight, which is coded according to the following system.

-   `carrier`: Two letter carrier abbreviation.

    -   `9E`: Endeavor Air Inc.
    -   `AA`: American Airlines Inc.
    -   `AS`: Alaska Airlines Inc.
    -   `B6`: JetBlue Airways
    -   `DL`: Delta Air Lines Inc.
    -   `EV`: ExpressJet Airlines Inc.
    -   `F9`: Frontier Airlines Inc.
    -   `FL`: AirTran Airways Corporation
    -   `HA`: Hawaiian Airlines Inc.
    -   `MQ`: Envoy Air
    -   `OO`: SkyWest Airlines Inc.
    -   `UA`: United Air Lines Inc.
    -   `US`: US Airways Inc.
    -   `VX`: Virgin America
    -   `WN`: Southwest Airlines Co.
    -   `YV`: Mesa Airlines Inc.

Remember that you can use `glimpse` to take a quick peek at your data to understand its contents better.

```{r}
glimpse(nycflights)
```

1. (a) How large is the `nycflights` dataset? (i.e. How many rows and columns does it have?)
<!-- Your response goes below this line! -->
Rows: 32,735
Columns: 16

<div style="color:blue">
Correct! 
</div>


1. (b) When were these data collected? 
<!-- Your response goes below this line! -->
2013

<div style="color:blue">
Correct! Did you use the help file? 
</div>


1. (c) Are there categorical variables in the dataset? If so, what are their names and data types? 
<!-- Your response goes below this line! -->

Carrier: chr
Tailnum: chr
Origin: chr
Dest: chr

<div style="color:blue">
Nice work! Keep in mind that if these variables were labeled a `fct` that would *also* be a categorical variable!  
</div>

<div style="color:blue">
Overall: 4
</div>


### Departure Delays

Let's start by examining the distribution of departure delays of all flights with a histogram.

```{r}
ggplot(data = nycflights, aes(x = dep_delay)) +
  geom_histogram()
```

This function says to plot the `dep_delay` variable from the `nycflights` data frame on the x-axis.
It also defines a `geom` (short for geometric object), which describes the type of plot you will produce.

Histograms are generally a very good way to see the shape of a single distribution of numerical data, but that shape can change depending on how the data is split between the different bins.
You can easily define the binwidth you want to use, by specifying the `binwidth` argument to `geom_histogram()`, like so:

```{r}
ggplot(data = nycflights, aes(x = dep_delay)) +
  geom_histogram(binwidth = 15)

ggplot(data = nycflights, aes(x = dep_delay)) +
  geom_histogram(binwidth = 150)
```

2.  Look carefully at these three histograms. How do they compare? Are features revealed in one that are obscured in another?
<!-- Your answer goes below this line -- do not remove this!  -->
All three of these histograms show us that a large majority of flights have a small delay. Looking at the last histogram with the largest bin size, one would conclude that a flight is likely going to have no delay or a very small delay, but a small number of flights experience some delay. The Histogram with the smallest binning size shows us that a majority of flights depart on time, and there is a good chance that your flight might depart early, or be slightly delayed.

<div style="color:blue">
Nice work! The smaller binwidth makes certain aspects of plot more visible. Do you think there would be a binwidth that would be too small?   
</div>

<div style="color:blue">
Overall: 4
</div>

3. A histogram is only one way to visualize the distribution of a numerical variable. 
(a) What are other visualizations that we could have produced? 
(b) What are the benefits and weaknesses of a histogram compared to these other options? 
<!-- Your response goes below this line! -->
We could've produced a dotplot, a boxplot, or a density plot.

A *dotplot* has the advantage of showing us more precisely when the flights are actually leaving and does not bin the data. As a histogram decreases its bin size, it approaches a dot plot. The dotplots are generally used with smaller datasets and the large number of flights around 0 may make the chart hard to read.

A *boxplot* can show us the outliers in a dataset and where the median, 25th, and 75th percentile lie. This can be helpful to understand what the general spread of the data is without giving exact counts. The downside is that for a dataset like this one where the average is clustered around a single value (0 in this case) all the other outliers will must be shown on the plot and will distort the the "box". The data shown in this "box" is generally why one would use a boxplot so for this dataset it is ill-advised. The only real advantage would be showing where the outliers in the dataset are.

A *density plot* is similar to a histogram, but does not give exact counts of data. For a dataset like this one, a density plot would have a massive spike at the beginning than flatten out very quickly. This results in a massively right skewed plot that gives one less information than the other plots, as it is hard to tell how many flights are actually being delayed and not a result of the curve encompassing the entire range of values.

<div style="color:blue">
Great description of the components and benefits of the different plotting options! 
</div>

<div style="color:blue">
Overall: 4
</div>


## LAX Destinations 

If you want to visualize only on delays of flights headed to Los Angeles, you need to first `filter()` the data for flights with that destination (`dest == "LAX"`) and then make a histogram of the departure delays of only those flights.

```{r}
lax_flights <- nycflights %>%
  filter(dest == "LAX")

ggplot(data = lax_flights, aes(x = dep_delay)) +
  geom_histogram()
```

Let's decipher these two commands (OK, so it might look like four lines, but the first two physical lines of code are actually part of the same command. It's common to add a break to a new line after `%>%` to help readability).

-   Command 1: Take the `nycflights` data frame, `filter()` for flights headed to LAX, and save the result as a new data frame called `lax_flights`.

    -   `==` means "if it's equal to".
    -   `LAX` is in quotation marks since it is a character string.

-   Command 2: Basically the same `ggplot()` call from earlier for making a histogram, except that it uses the smaller data frame for flights headed to LAX instead of all flights.

**Logical operators:** Filtering for certain observations (e.g. flights from a particular airport) is often of interest in data frames where we might want to examine observations with certain characteristics separately from the rest of the data.
To do so, you can use the `filter()` function and a series of **logical operators**.
The most commonly used logical operators for data analysis are as follows:

-   `==` means "equal to"
-   `!=` means "not equal to"
-   `>` or `<` means "greater than" or "less than"
-   `>=` or `<=` means "greater than or equal to" or "less than or equal to"


4. You learned in the tutorial that this process of creating a new dataset based on a filter is not necessarily the best plan. An alternative plan is to use the pipe operator (`%>%`) to create a data pipeline, carrying the filtered data into the plot. 
Change the code below, so it uses the pipe operator instead of saving a new dataset. 

```{r}
## For this exercise, you need to change the code below! 

nycflights %>%
  filter(dest == "LAX") %>%

ggplot(aes(x = dep_delay)) +
  geom_histogram()
## having a blank line between the two steps feels a bit awkward and could easily be interpreted as two different components!
```

<div style="color:blue">
Great use of the pipe operator!
See my comments in your code above! 
</div>

<div style="color:blue">
Overall: 4
</div>



### Multiple Data Filters 

You can also filter based on multiple criteria.
Suppose you are interested in flights headed to San Francisco (SFO) in February:

```{r}
sfo_feb_flights <- nycflights %>%
  filter(dest == "SFO", month == 2)
```

Note that you can separate the conditions using commas if you want flights that are both headed to SFO **and** in February.
If you are interested in either flights headed to SFO **or** in February, you can use the `|` instead of the comma.


5.  Find the number of flights in this dataset that were headed to SFO in February. 
**Hint:** The `filter()` and `count()` functions will be helpful!  

```{r}
## Code for exercise 1 here! 
nycflights %>%
  filter(dest=="SFO" & month==2) %>%
  ## I'd recommend adding spaces before and after your = signs in R, to improve your code readability. 
  ## Also, remember that the comma in filter() does the same thing as an & operator. 
  count()

```

<div style="color:blue">
Nice work!
See my comments in your code above. 
</div>

<div style="color:blue">
Overall: 4
</div>

## Data Summaries 

You can also obtain numerical summaries for the flights headed to LAX, using the `summarise()` function:

```{r}
lax_flights %>%
  summarise(mean_dd   = mean(dep_delay), 
            median_dd = median(dep_delay), 
            n         = n())
```

Note that in the `summarise()` function you created a list of three different numerical summaries that you were interested in.
The names of these elements are user defined, like `mean_dd`, `median_dd`, `n`, and you can customize these names as you like (just don't use spaces in your names).
Calculating these summary statistics also requires that you know the summary functions you would like to use. 

**Summary statistics:** Some useful function calls for summary statistics for a single numerical variable are as follows:

-   `mean()`: calculates the average
-   `median()`: calculates the median
-   `sd()`: calculates the standard deviation
-   `var()`: calculates the variances
-   `IQR()`: calculates the inner quartile range (Q3 - Q1)
-   `min()`: finds the minimum 
-   `max()`: finds the maximum
-   `n()`: reports the sample size

Note that each of these functions takes a single vector as an argument and returns a single value.

## Data Summaries & Data Visualizations 

6.  Describe the distribution of the **arrival** delays of the LAX flights using a histogram and appropriate summary statistics.
**Hint:** The summary statistics you use should depend on the shape of the distribution.
<!-- Your response to this problem goes below this line! -->

<div style="color:blue">
Great visualization and data summaries! Are the summaries you found appropriate for this distribution?  
Where is your description? 
</div>

<div style="color:blue">
Overall: Redo
</div>


```{r}
## Code for exercise 2 goes here! 
lax_flights <- nycflights %>%
  filter(dest=="LAX")
ggplot(lax_flights, aes(x=arr_delay)) + 
  geom_histogram()

lax_flights %>%
  summarise(median_arrival_delay = median(arr_delay),
            IQR_arrival_delay = IQR(arr_delay),
            min_arrival_delay = min(arr_delay),
            max_arrival_delay = max(arr_delay),
            n = n())
```

## Grouped Summaries 

Another useful technique is quickly calculating summary statistics for various groups in your data frame.
For example, we can modify the above command using the `group_by()` function to get the same summary stats for each origin airport:

```{r}
sfo_feb_flights %>%
  group_by(origin) %>%
  summarise(median_dd = median(dep_delay), 
            iqr_dd = IQR(dep_delay), 
            n_flights = n())
```

Here, we first grouped the data by `origin` and then calculated the summary statistics.

## Summaries vs. Visualizations 

Which month would you expect to have the highest average delay departing from an NYC airport?

Let's think about how you could answer this question. 
One option is to summarize the data and inspect the output. 
Another option is to plot the delays and inspect the plots. 
Let's try both! 

7. Find the average departure delays for each month, and arrange these average delays in descending order. 
**Hint:** the `desc()` function is helpful for arranging numbers in descending order! 

```{r}
## Code for exercise 3 goes here! 
nycflights %>%
  group_by(month) %>%
  summarise(mean_dd = mean(dep_delay),
            median_dd = median(dep_delay))

```

<div style="color:blue">
You're almost there! You've found the data summaries, but now you need to arrange them in descending order! 
</div>

<div style="color:blue">
Overall: Redo
</div>


8. Now, plot the distribution of departure delays, faceted by the month of departure. 
Choose the type of plot you believe is appropriate for visualizing the distribution of departure delays. 

```{r}
## Code for exercise 4 goes here! 
nycflights %>%
  group_by(month) %>%
  ggplot(aes(x=dep_delay)) +
    geom_histogram(binwidth=20) +
    facet_wrap(vars(month))

```

<div style="color:blue">
Nice work faceting by month! You do not need to `group_by()` variables before you plot, as `ggplot()` ungroups the data before plotting. Coloring or faceting is similar to grouping the data when plotting.  
</div>

<div style="color:blue">
Overall: 4
</div>



9. What information can you obtain from the visualization that you could not from the data summaries?   
<!-- Your response to this question goes below this line!  -->


One piece of data that can be extracted from the graphs and not the summary, is the quantity of flights that are on time or nearly so. Looking at the graphn it becomes apparent that the number of on time flights dwarfs the flights with any significant delay.
