---
title: "Introduction to Visualizing and Summarizing Data"
author: "Shannon Maguire"
date: "Due: January 24, 2021 at 11:59pm"
output: html_document
---

<div style="color:blue">

Great job! You have revisions on problems: 

My comments for your responses to the exercises can be found in blue text, and my comments/suggestions for your R code can be found in the code chunks behind two hashtags (##)

Your homework assignment receives a score that reflects the average points-per-question, where each question is worth 4 points. This is assigned based on the grading system outlined in the syllabus. 

Problems marked with an “R” must be redone and returned within two days of receiving the assignment back. You will receive a score of 0 on each problem requiring a revision until a revision is submitted. Late revisions receive a zero, unless you contact me. Work successfully redone (on time) will receive a maximum grade of 3. It is possible to have a redo on a redo.

</div>

```{r, include = FALSE}
knitr::opts_chunk$set(echol = TRUE, 
                      results = "hold", 
                      message = FALSE, 
                      warning = FALSE)
```

## Getting started

### Load packages

In this lab, we will explore and visualize the data using the **tidyverse** suite of packages. The data can be found in the companion package for OpenIntro labs, **openintro**.

Let's load the following packages:

-   The **tidyverse** "umbrella" package which houses a suite of many different `R` packages: for data wrangling and data visualization
-   The **openintro** `R` package: for data and custom functions with the OpenIntro resources

```{r, include = FALSE}
library(tidyverse)
library(openintro)
## This is a comment -- R will not run this line as code, because it has ## before it
```

### The data

The [Bureau of Transportation Statistics](http://www.rita.dot.gov/bts/about/) (BTS) is a statistical agency that is a part of the Research and Innovative Technology Administration (RITA).
As its name implies, BTS collects and makes transportation data available, such as the flights data we will be working with in this lab.

First, we'll view the `nycflights` data frame.
Type the following in your console to load the data:

```{r}
data(nycflights)
```

The data set `nycflights` that shows up in your workspace is a *data matrix*, with each row representing an *observation* and each column representing a *variable*.
R calls this data format a **data frame**, which is a term that will be used throughout the labs.
For this data set, each *observation* is a single flight.

To view the names of the variables, you can use the `names()` function: 

```{r}
names(nycflights)
```

This returns the names of the variables in this data frame.
The **codebook** (description of the variables) can be accessed by pulling up the help file by typing a `?` before the name of the dataset:

```{r, eval = FALSE}
?nycflights
```

One of the variables refers to the carrier (i.e. airline) of the flight, which is coded according to the following system.

-   `carrier`: Two letter carrier abbreviation.

    -   `9E`: Endeavor Air Inc.
    -   `AA`: American Airlines Inc.
    -   `AS`: Alaska Airlines Inc.
    -   `B6`: JetBlue Airways
    -   `DL`: Delta Air Lines Inc.
    -   `EV`: ExpressJet Airlines Inc.
    -   `F9`: Frontier Airlines Inc.
    -   `FL`: AirTran Airways Corporation
    -   `HA`: Hawaiian Airlines Inc.
    -   `MQ`: Envoy Air
    -   `OO`: SkyWest Airlines Inc.
    -   `UA`: United Air Lines Inc.
    -   `US`: US Airways Inc.
    -   `VX`: Virgin America
    -   `WN`: Southwest Airlines Co.
    -   `YV`: Mesa Airlines Inc.

Remember that you can use `glimpse` to take a quick peek at your data to understand its contents better.

```{r}
glimpse(nycflights)
```

1. (a) How large is the `nycflights` dataset? (i.e. How many rows and columns does it have?)
<!-- Your response goes below this line! -->
There are 32,735 rows and 16 columns in the dataset.


1. (b) When were these data collected? 
<!-- Your response goes below this line! -->
This data was collected in 2013.


1. (c) Are there categorical variables in the dataset? If so, what are their names and data types? 
<!-- Your response goes below this line! -->
Yes, the variables carrier, tailnum, flight, origin, and dest are categorical variables. Flight is an integer variable, while the other variables are character variables.


### Departure Delays

Let's start by examining the distribution of departure delays of all flights with a histogram.

```{r}
ggplot(data = nycflights, aes(x = dep_delay)) +
  geom_histogram()
```

This function says to plot the `dep_delay` variable from the `nycflights` data frame on the x-axis.
It also defines a `geom` (short for geometric object), which describes the type of plot you will produce.

Histograms are generally a very good way to see the shape of a single distribution of numerical data, but that shape can change depending on how the data is split between the different bins.
You can easily define the binwidth you want to use, by specifying the `binwidth` argument to `geom_histogram()`, like so:

```{r}
ggplot(data = nycflights, aes(x = dep_delay)) +
  geom_histogram(binwidth = 15)

ggplot(data = nycflights, aes(x = dep_delay)) +
  geom_histogram(binwidth = 150)
```

2.  Look carefully at these three histograms. How do they compare? Are features revealed in one that are obscured in another?
<!-- Your answer goes below this line -- do not remove this!  -->
The histograms with the highest number of bins (lowest bin width) give us the most specific detail, making it possible to pinpoint the specific range where the most common departure delay times occur. Additionally, in the second histogram you are able to see how many departure delays were negative (meaning that the flights arrived early), which you can't see from the first and third histograms since they are grouped with some positive departure delay times.

3. A histogram is only one way to visualize the distribution of a numerical variable. 
(a) What are other visualizations that we could have produced? 
(b) What are the benefits and weaknesses of a histogram compared to these other options? 
<!-- Your response goes below this line! -->
You could also use a boxplot or a stem-and-leaf plot to visualize the distribution of dep_delay. Histograms are useful because they are simple and very easy for anyone to digest. They are consistent in nature because the bins are all of equal width, which makes them easier to interpret. However, you don't quite get as much information from a histogram than you would a boxplot, since a boxplot can clearly show you things like median, quartiles, and outliers. Boxplots are also easier to compare side-by-side.


## LAX Destinations 

If you want to visualize only on delays of flights headed to Los Angeles, you need to first `filter()` the data for flights with that destination (`dest == "LAX"`) and then make a histogram of the departure delays of only those flights.

```{r}
lax_flights <- nycflights %>%
  filter(dest == "LAX")

ggplot(data = lax_flights, aes(x = dep_delay)) +
  geom_histogram()
```

Let's decipher these two commands (OK, so it might look like four lines, but the first two physical lines of code are actually part of the same command. It's common to add a break to a new line after `%>%` to help readability).

-   Command 1: Take the `nycflights` data frame, `filter()` for flights headed to LAX, and save the result as a new data frame called `lax_flights`.

    -   `==` means "if it's equal to".
    -   `LAX` is in quotation marks since it is a character string.

-   Command 2: Basically the same `ggplot()` call from earlier for making a histogram, except that it uses the smaller data frame for flights headed to LAX instead of all flights.

**Logical operators:** Filtering for certain observations (e.g. flights from a particular airport) is often of interest in data frames where we might want to examine observations with certain characteristics separately from the rest of the data.
To do so, you can use the `filter()` function and a series of **logical operators**.
The most commonly used logical operators for data analysis are as follows:

-   `==` means "equal to"
-   `!=` means "not equal to"
-   `>` or `<` means "greater than" or "less than"
-   `>=` or `<=` means "greater than or equal to" or "less than or equal to"


4. You learned in the tutorial that this process of creating a new dataset based on a filter is not necessarily the best plan. An alternative plan is to use the pipe operator (`%>%`) to create a data pipeline, carrying the filtered data into the plot. 
Change the code below, so it uses the pipe operator instead of saving a new dataset. 

```{r}
## For this exercise, you need to change the code below! 

nycflights %>%
  filter(dest == "LAX") %>%
ggplot(aes(x = dep_delay)) +
    geom_histogram()
```


### Multiple Data Filters 

You can also filter based on multiple criteria.
Suppose you are interested in flights headed to San Francisco (SFO) in February:

```{r}
sfo_feb_flights <- nycflights %>%
  filter(dest == "SFO", month == 2)
```

Note that you can separate the conditions using commas if you want flights that are both headed to SFO **and** in February.
If you are interested in either flights headed to SFO **or** in February, you can use the `|` instead of the comma.


5.  Find the number of flights in this dataset that were headed to SFO in February. 
**Hint:** The `filter()` and `count()` functions will be helpful!  

```{r}
## Code for exercise 1 here! 

nycflights %>%
  filter(dest == "SFO", month == 2) %>%
  count()
```


## Data Summaries 

You can also obtain numerical summaries for the flights headed to LAX, using the `summarise()` function:

```{r}
lax_flights %>%
  summarise(mean_dd   = mean(dep_delay), 
            median_dd = median(dep_delay), 
            n         = n())
```

Note that in the `summarise()` function you created a list of three different numerical summaries that you were interested in.
The names of these elements are user defined, like `mean_dd`, `median_dd`, `n`, and you can customize these names as you like (just don't use spaces in your names).
Calculating these summary statistics also requires that you know the summary functions you would like to use. 

**Summary statistics:** Some useful function calls for summary statistics for a single numerical variable are as follows:

-   `mean()`: calculates the average
-   `median()`: calculates the median
-   `sd()`: calculates the standard deviation
-   `var()`: calculates the variances
-   `IQR()`: calculates the inner quartile range (Q3 - Q1)
-   `min()`: finds the minimum 
-   `max()`: finds the maximum
-   `n()`: reports the sample size

Note that each of these functions takes a single vector as an argument and returns a single value.

## Data Summaries & Data Visualizations 

6.  Describe the distribution of the **arrival** delays of the LAX flights using a histogram and appropriate summary statistics.
**Hint:** The summary statistics you use should depend on the shape of the distribution.
<!-- Your response to this problem goes below this line! -->
Since the distribution of arrival delays is skewed to the right, we will use median and IQR instead of mean and SD. For the 1,583 LAX flights, the median arrival delay is 6 minutes early and the IQR is 30.5 minutes. The earliest a flight has ever been is 62 minutes early, and the latest a flight has ever been is 319 minutes late.


```{r}
## Code for exercise 2 goes here! 

ggplot(data = lax_flights, aes(x = arr_delay)) +
  geom_histogram()

lax_flights %>%
  summarise(median_ad = median(arr_delay),
            iqr_ad    = IQR(arr_delay),
            min_ad    = min(arr_delay),
            max_ad    = max(arr_delay),
            n         = n())
```

## Grouped Summaries 

Another useful technique is quickly calculating summary statistics for various groups in your data frame.
For example, we can modify the above command using the `group_by()` function to get the same summary stats for each origin airport:

```{r}
sfo_feb_flights %>%
  group_by(origin) %>%
  summarise(median_dd = median(dep_delay), 
            iqr_dd = IQR(dep_delay), 
            n_flights = n())
```

Here, we first grouped the data by `origin` and then calculated the summary statistics.

## Summaries vs. Visualizations 

Which month would you expect to have the highest average delay departing from an NYC airport?

Let's think about how you could answer this question. 
One option is to summarize the data and inspect the output. 
Another option is to plot the delays and inspect the plots. 
Let's try both! 

7. Find the average departure delays for each month, and arrange these average delays in descending order. 
**Hint:** the `desc()` function is helpful for arranging numbers in descending order! 

```{r}
## Code for exercise 3 goes here! 

nycflights %>%
  group_by(month) %>%
  summarise(mean_dd = mean(dep_delay)) %>%
  arrange(desc(mean_dd))
```

8. Now, plot the distribution of departure delays, faceted by the month of departure. 
Choose the type of plot you believe is appropriate for visualizing the distribution of departure delays. 

```{r}
## Code for exercise 4 goes here! 

nycflights %>%
  group_by(month) %>%
  ggplot(aes(y = dep_delay)) +
  geom_boxplot() +
  facet_wrap(~month, nrow = 1) +
  ylim(-30, 500) # This limit excludes some data but makes it easier to compare the months
```

9. What information can you obtain from the visualization that you could not from the data summaries?   
<!-- Your response to this question goes below this line!  -->
Looking at the visualization, you can more easily tell which months had more variation by seeing with boxes are taller. You can also see specific points that are outliers for each month, and are able to better distinguish and compare the shape of the distributions for each month

