---
title: "Summarizing Numerical Data"
date: "October 5, 2021"
author: "Dr. Allison Theobold"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "slide-style.css"]
    nature:
      highlightStyle: solarized-light
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---


```{r, echo = FALSE, message = FALSE, warning = FALSE}
# R options
options(
  htmltools.dir.version = FALSE,
  tibble.width = 65,
  width = 65
  )

# figure height, width, dpi
knitr::opts_chunk$set(fig.width = 8, 
                      fig.asp = 0.618,
                      out.width = "60%",
                      dpi = 300, 
                      warning = FALSE, 
                      message = FALSE)

# fontawesome
htmltools::tagList(rmarkdown::html_dependency_font_awesome())

# magick
dev.off <- function(){
  invisible(grDevices::dev.off())
}

# xaringanExtra
library(xaringanExtra)
xaringanExtra::use_panelset()

library(emo)
library(palmerpenguins)
library(tidyverse)
library(openintro)
library(ggridges)
library(flair)
library(scales)
library(gghighlight)
```

```{r set-theme, include = FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color      = "#b76352", # mango
  secondary_color    = "#34605f", # bayberry
  header_font_google = google_font("Raleway"),
  text_font_google   = google_font("Raleway", "300", "300i"),
  code_font_google   = google_font("Source Code Pro"),
  header_color = "#793540", #rhubarb
  white_color = "#F5F5F5", # lightest color
  black_color = "#36454F", # darkest color
  text_font_size = "30px", 
  link_color = "#696969" #grey
)
```

class: center, middle, inverse 

Given this distribution... 

```{r, echo = FALSE}
penguins %>% 
  ggplot(aes(x = body_mass_g)) + 
  geom_histogram(binwidth = 200) + 
  labs(x = "Body Mass (grams)")
```

.large[What measure of center would you use? Why?]

---

class: center, middle 

.huge[Summary Statistics]

--

.bayberry[.larger[Measures of Center]]

.mango[.larger[Measures of Spread]]

---

class: middle 

.bayberry[.larger[Measures of Center]]

.pull-left[
Mean
  - average of data values
  - not resistant to outliers
]

--

.pull-right[
Median
  - middle observation 
  - resistant to outliers
]

---

class: center, middle, inverse

.larger[For fairly normal data...]

```{r, echo = FALSE}

x_bar <- mean(penguins$body_mass_g, na.rm = TRUE)
m <- median(penguins$body_mass_g, na.rm = TRUE)

penguins %>% 
  ggplot(aes(x = body_mass_g)) + 
  geom_histogram(binwidth = 200) + 
  geom_vline(xintercept = x_bar, color = "red", lwd = 1) +
  geom_vline(xintercept = m, color = "blue", lwd = 1) +
  geom_text(x = 3800, y = 20, label = "Median", color = "blue", size = 5) + 
  geom_text(x = 4400, y = 15, label = "Mean", color = "red", size = 5) + 
  labs(x = "Body Mass (grams)")
```

---

class: center, middle, inverse 

.larger[For bimodal data...]

```{r, echo = FALSE}

x_bar <- mean(penguins$bill_length_mm, na.rm = TRUE)
m <- median(penguins$bill_length_mm, na.rm = TRUE)

penguins %>% 
  ggplot(aes(x = bill_length_mm)) + 
  geom_histogram(binwidth = 2) + 
  geom_vline(xintercept = x_bar, color = "red", lwd = 1) +
  geom_vline(xintercept = m, color = "blue", lwd = 1) +
  geom_text(x = 47, y = 20, label = "Median", color = "blue", size = 5) + 
  geom_text(x = 42, y = 15, label = "Mean", color = "red", size = 5) + 
  labs(x = "Bill Length (grams)")
```

---

class: middle

.mango[.larger[Measures of Spread]]

.pull-left[
- Variance
  * average squared distance from the mean
  * not resistant to outliers

- Range
  * difference between minimum and maximum
  * not resistant to outliers
  
]

--

.pull-right[

</br>
</br>

- Inner Quartile Range (IQR) 

  * difference between Q1 and Q3
  * resistant to outliers
]  
  
---

.huge[.hand[Your Turn]]

.larger[What are two benefits and two weaknesses of data summaries and data 
visualizations?]

---

class: middle, center, inverse

.larger[Point Estimates & Parameters]

*Parameter*: True value of the statistic for the population of interest  

--
  
*Point Estimate*: provides our best guess for the value of the parameter 

--

Estimates based on larger samples tend to be more accurate than those based on smaller samples.


---

.larger[.hand[Your Turn]]

.pull-left-wide[
Describe what you would need to do to replicate the following plot.

```{r, echo = FALSE}
nycflights %>%
  filter(carrier %in% c("AA", "UA", "DL")) %>% 
  ggplot(aes(x = dep_delay, y = arr_delay, color = carrier)) + 
  geom_point()
```

]

--

.pull-right-narrow[

Next...

What are two aspects of this plot that you believe are well done? What are 
two aspects of the plot that could be improved?  
]

