---
title: "Setting Up R and RStudio"
output:
  html_document:
    df_print: paged
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: |
  Learn how to set up R and RStudio on your computer. We will also demonstrate how to install R packages from CRAN, and install the tidyverse package (a package we will use often).
---


```{r setup, include=FALSE}
library(learnr)
tutorial_options(exercise.timelimit = 60)
```

## Welcome

This is a demo tutorial that was made using an RMarkdown file and compiled 
(made) into an HTML file. If you are interested, 
you can inspect the [source code](https://github.com/rstudio/learnr/tree/master/inst/tutorials/ex-setup-r/ex-setup-r.Rmd)
that was used to made it. We will be using RMarkdown files later on in the 
course!

### Outline

This tutorial will help you set up your computer to use R. It is for you if you
need to:

* Change where files are downloaded
* Install R on your computer
* Install the RStudio IDE
* Learn how to work in RStudio
* Install the `tidyverse` R package

You can skip this tutorial if you've already done these things.

### Is this tutorial for you?

Do you need to work through the tutorial? Take the quiz below to find out.

```{r quiz1, echo = FALSE}
question("Check all that you have NOT done:",
  answer("installed `R` on my computer", 
         message = "* Install `R`"),
  answer("installed the RStudio IDE",
         message = "* Install RStudio IDE"),
  answer("installed the tidyverse R package",
         message = "* Install Packages"),
  answer("None of the above. I've done them all.",
         correct = TRUE, 
         message = "You can skip this tutorial!"),
  type = "multiple",
  incorrect = "This tutorial is here to help! To get set up read:"
)
```

## Change File Download Destination

I would highly recommended you change the default file download location for
your internet browser!

Generally by default, internet browsers automatically save all files to the
Downloads folder on your computer. This does not encourage good file
organization practices. It is highly recommended that you change this option so 
that your browser asks you where to save each file before downloading it.

This [page](https://www.online-tech-tips.com/computer-tips/change-default-download-folder-location-on-any-web-browser/) has information on how to do this for the most common browsers.


## Install `R`

### How to install `R`

![](https://vimeo.com/203516510)

The video was recorded with a previous version of `R`. The current version of
`R` is:

```{r, echo = FALSE}
R.version$version.string
```

### Test your knowledge

```{r quiz2_1, echo=FALSE}
question("Is `R` free to download and use?",
    answer("Yes!", correct = TRUE, 
           message = "`R` is free _and_ open source, which means that anyone can
           read, use, and copy the source code of the `R` language."),
    answer("No."), 
    allow_retry = TRUE
    )
```

```{r quiz2_2, echo = FALSE}
question("Where do you download `R`?",
  answer("www.rstudio.com/download",
         message = "This is where you download RStudio, not `R`."),
  answer("[cloud.r-project.org](http://cloud.r-project.org)", 
         correct = TRUE,
         message = "You can also download R from
         [cran.r-project.org](http://cran.r-project.org)"),
  answer("www.r-project.org", 
         message = "Good try, but not exactly. www.r-project.org doesn't 
         provide a download link, but it does provide a link to one the 
         websites above."),
  answer("www.r.com", 
         message = "Not quite, [dot]com is reserved for businesses. `R` is a
         non-profit organization, so it uses a [dot]org web address."),
  allow_retry = TRUE
  )
```

```{r quiz2_3, echo = FALSE}
  question("How often should you update `R`?",
    answer("Everytime you use it",
           message = "This will be too often unless you use `R` very rarely!"),
    answer("About once a year", 
           correct = TRUE, 
           message = "A new version of `R` is released about once a year. Update
           sooner if you encounter a bug that you cannot explain."),
    answer("Never",
           message = "A new version of `R` is released about once a year. In our 
           course, I'll assume that you are using the newest version of `R`,
           which will be the fastest version with the fewest unexpected
           behaviors."),
    allow_retry = TRUE
  )
```

## Install RStudio

### How to install RStudio

RStudio is an Integrated Development Environment (IDE) for `R`. What does that
mean? Well, if you think of `R` as a language, which it is, you can think of
RStudio as a program that helps you write and work in the language. RStudio
makes programming in `R` much easier and I suggest that you use it!

![](https://vimeo.com/203516968)

### Test your knowledge

```{r quiz3, echo=FALSE}
quiz(caption = "Quiz - Install RStudio",
  question("What is the RStudio IDE?",
    answer("An application that makes it easier to use `R`.",
           correct = TRUE, 
           message = "RStudio organizes your workspace and makes it easier to 
           write, use, debug, and save `R` code. I highly recommend it and will 
           rely on the RStudio IDE in several tutorials."),
    answer("An application that let's you use R without writing any code.",
           message = "And thank goodness! Code provides a reproducible record of
           your work, which is essential for data science."),
    answer("A spreadsheet program like Microsoft Excel.", 
           message = "RStudio doesn't allow for you to manipulate Excel files 
           directly, instead you have to write some code to work with your
           data!"),
    answer("Another name for `R`", 
           message = "`R` and RStudio are two separate things. `R` is a
           language, like English. Think of RStudio as a program that helps you
           use the language, kind of like how a word processing program helps 
           you write in English."),
    allow_retry = TRUE
  ),
  
  question("Is the RStudio IDE free to download and use?",
    answer("Yes!",
           correct = TRUE,
           message = "Like `R`, RStudio is free and open-source. There do exist
           professional versions of RStudio with enhanced features, but we will
           not rely on those features in these tutorials."),
    answer("No.",
           message = "Like `R`, RStudio is free and open-source. There do exist 
           professional versions of RStudio with enhanced features, but we will
           not rely on those features in these tutorials."), 
    allow_retry = TRUE
  ),
  
  question("Where do you download RStudio?",
    answer("www.rstudio.com/download",
           correct = TRUE, 
           message = "For these tutorials, download and install the RStudio
           Desktop Open Source License."),
    answer("[cloud.r-project.org](http://cloud.r-project.org)",
           message = "This is where you download `R`, not RStudio."),
    answer("www.r-project.org"),
    answer("[cran.rstudio.org](http://cran.rstudio.org)"),
    allow_retry = TRUE
  ),
  
  question("Do you need to install `R` if you already have RStudio?",
    answer("Yes.",
           correct = TRUE),
    answer("No.",
           message = "RStudio does not come with `R`, but uses `R` to excute 
           (run) the code that you write in RStudio. You need to install `R`
           separately to be able to run code in RStudio.")
  )
)
```

## How to Work in RStudio, Save Files, and Open Downloaded Files

Please watch this video made by Professor Lisa Lendway that describes essential
configuration options for RStudio.

![](https://www.youtube.com/watch?v=LJiRSPoe8r0)

## Install Packages

### How to install R packages

![](http://youtu.be/tXZDoMlXfTE?hd=1)

### Test your knowledge

```{r names, echo = FALSE}
quiz(caption = "Quiz - Working with Packages",
  question("What are the two way you could install packages in R?",
    answer("Run `library()` with the package you want to install.",
           message = "Not quite! `library()` is the function we will use to 
           load a package into `R` that we have already installed."),
    answer("Run `install.packages()` with the name of the packages you want to 
           install.",
           correct = TRUE),
    answer("`install_packages()`", 
           message = "Careful! In `R` the punctuation of a function matters.
           There is no `install_packages()` function, only an
           `install.packages()` function."),
    answer("Click on the Install button in the Packages tab, type in the 
           name of the package you would like to install, and click Install.", 
           correct = TRUE),
    answer("There is no command. You must visit
           [cran.r-project.org](http://cran.r-project.org) and download packages
           manually.", 
           message = "Doing this everytime would be **extrememly** difficult!
           Instead, `R` makes it easy to download packages. You will need to
           connect to the internet and then run one of the commands above. Find
           the command!"),
    incorrect = "Make sure you find both ways!", 
    allow_retry = TRUE
  ),
  
  question("How often do you need to install a package on your computer?",
    answer("Every time you restart R", 
           message = "You only need to install a package once. You may need to 
           update the package intermettently, but the instalation only happens 
           once!"),
    answer("Every time you restart your computer", 
           message = "You only need to install a package once. You may need to 
           update the package intermettently, but the instalation only happens 
           once!"),
    answer("Only once. Afterwards, `R` can find the package on your hard drive
    when it needs to.",
           correct = TRUE),
    answer("Never, as long as you are connected to the internet.", 
           message = "This could be true if you are using `R` over a cloud
           service. However, for this class we will be using either a local 
           if you are using `R` locally on your own computer,
           you will need to install each package that you use on your
           computer."),
    allow_retry = TRUE
  ),
  
  question("What is the tidyverse?",
    answer("A collection of packages that work well together and provide tools
           for common data science tasks.",
           correct = TRUE,
           message = 'Correct! You can install all of the tidyverse packages at
           once with `install.packages("tidyverse")`.'),
    answer("A word processing program for `R`", 
           message = "While RStudio could be sort of thought of as a word 
           processing program for writing `R` code, the `tidyverse` lives inside 
           of `R`." ),
    answer("A code style that reads like poetry", 
           message = "I agree that the language of the `tidyverse` reads 
           incredibly well! But, that's not **exactly** what the 
           `tidyverse` is."),
    answer("An alternate universe where everyone is an `R` programmer.", 
           message = "Hell, I really wish this was true!"),
    allow_retry = TRUE
  )
)
```

## Troubleshooting 

Problem: You are on a Mac and getting the following error:

```
Error: package or namespace load failed for ‘tidyverse’ in loadNamespace(i, c(lib.loc, .libPaths()), versionCheck = vI[[i]]):
 there is no package called ‘rlang’`
```

Here’s how to fix it:

- First install the suite of Command Line Tools for Mac using the instructions (here)[https://osxdaily.com/2014/02/12/install-command-line-tools-mac-os-x/]. 

- Next enter install.packages("rlang") in the Console. 

- Finally check that entering library(ggplot2) gives no errors. You should be
good now for tidyverse and openintro.