---
title: "Week Ten: Inference for Many Means (ANOVA)"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
library(emo)
library(gridExtra)
```


Welcome!

In this week's coursework we are wrapping up the quarter by investigating how 
we can use the inference we've learned in a new context---comparing multiple 
means. These comparisons have a specific name, ANalysis of VAriance (ANOVA). 


We are going to use all of the simulation-based methods we learned previously 
for this new context. An ANOVA relies on a new statistic, the $F$-statistic, to 
summarize how different 3 or more means are from each other. 

The primary focus of an ANOVA is to detect if the means of 3 or more groups are 
different. Because of this we focus on hypothesis tests for a difference in the 
group means and **do not** use confidence intervals. We will generate
permutation distributions of $F$-statistics that we could have observed
**if the null hypothesis was true** (there is no difference in the group means).


## Learning Outcomes 

By the end of this coursework you should be able to:

- outline what hypotheses we use to test for a linear relationship
- describe how a permutation distribution for a slope statistic is made
  * connect the null hypothesis with how observations are permuted
- describe how a bootstrap distribution for a slope statistic is made
- interpret a p-value in the context of a linear relationship 
- interpret a confidence interval in the context of a linear relationship
- explain the similarities and differences between parametric ($t$-based)
methods and non-parametric (simulation-based) methods


- use R to:
  * generate a permutation distribution for a slope
  * visualize the permutation distribution
  * calculate the observed slope statistic
  * calculate a p-value for a hypothesis test


---

## This Week's TODOs

`r emo::ji("book")` Reading: 45-60 minutes 

`r emo:ji("computer")` Tutorial: 30-45 minutes

`r emo::ji("paper")` Group Discussion Homework: ANOVA Project  

`r emo::ji("check")` Check-ins: 1

---

## Comparing Multiple Means

`r emo::ji("book")` [**Required Reading:** *IMS*, Inference for Comparing Many Means](https://openintro-ims.netlify.app/inference-many-means.html)

---

## 22.1 -- Case Study


---

## 22.2 -- Randomization Tests for Comparing Many Means

`r emo::ji("computer")` [R Tutorial](https://openintro.shinyapps.io/ims-05-infer-08/)

---

## 22.3 -- Mathematical Model for Comparing Many Means


---

## 22.4 -- Chapter Review 


---

## Group Project

**Details will be announced in class on Monday and posted here afterward.** 

---

## Think Out Loud



