---
title: "Week Two: Summarizing & Visualizing Categorical Data"
author: ""
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
library(flair)
library(emo)

```


Welcome!

In this week's coursework we are going to dive into exploring data, through 
data summaries and visualizations. We will be using `R` to carry out both of 
these tasks. We will start with the foundations of working in `R`, using 
data and installing / loading packages. Next, we will learn tools for
summarizing and visualizing categorical variables. 


## Learning Outcomes 

By the end of this coursework you should be able to:

- describe what a categorical variable is and how these types of variables 
are stored in `R` 
- determine whether a variable in a dataset is categorical
- create categorical variables from numerical variables 
- summarize categorical variables with frequencies of the different levels
- create visualizations of one and two categorical variables using side-by-side 
or stacked barcharts or facets 



## This Week's TODOs

`r emo::ji("book")` Readings: 30-45 minutes 

`r emo::ji("computer")` Tutorials: 60-75 minutes 

`r emo::ji("check")` Check-ins: 2 (submit on Canvas)

`r emo::ji("scroll")` Group Discussion Homework: 4 Questions

---

# Hello Data 

Let's start off by learning some of the fundamentals for working in `R`. If you 
are familiar with `R`, I would still encourage you to give this a try, in case 
some of the material is new to you! 

`r emo::ji("computer")` [**Required Tutorial:** Language of Data](https://openintro.shinyapps.io/ims-01-data-01)

Please work through __all__ of the topics _except_ Visualizing numerical data. 

</br> 

`r emo::ji("check")` **Check-in 1:** (Due Friday)

1. What function do you use to load in a package in `R`?

<!-- library() -->

2. What helpful function outputs the number of rows (observations), columns 
(variables), the names of the variables and their data types, and a preview of 
the first few entries of each variable?

<!-- glimpse() -->

3. A categorical variable could be **correctly** stored as what data types? 

- integer
- double
- character
- factor

4. What function do you use to retain only certain observations in 
a dataset (remove unwanted observations)? 

- `filter()`
- `count()`
- `mutate()`

5. What function did you learn that is used for converting a numerical variable
to a categorical variable (a process called discretizing)? 

<!-- if_else() -->

</br>

---

# Exploring Categorical Data 

`r emo::ji("book")` [**Required Reading:** Exploring Categorical Data](https://openintro-ims.netlify.app/explore-categorical.html)

`r emo::ji("computer")` [**Required Tutorial:** Exploratory Data Analysis, Exploring Categorical Data](https://openintro.shinyapps.io/ims-02-explore-01/)

<!-- `r emo::ji("play button")` **Supplementary Slides**  -->

<!-- - [Why use ggplot2?]() -->

<!-- - [First Steps with ggplot2]() -->

<!-- - [What is an **aes**thetic?]() -->

<!-- - [What is a **facet**?]() -->

<!-- - [Where do you want your bars?]() -->

<!-- - [What does `group_by()` do?]() -->

<!-- - [What does `pivot_wider()` do?]() -->

</br> 

`r emo::ji("scroll")` **Group Concept Discussion Questions** (Due Wednesday) 

1.	In the chapter readings and the tutorials you’ve learned about different
data types. Fill in the associated data types (e.g. character, factor, integer, 
double) with each type of variable. 
	
| Variable    | Data Type in R |
|-------------|----------------|
| Categorical variable |       |	
| Continuous numerical variable | |
| Discrete numerical variable	  | |


2.	Describe in your own terms what the pipe operator (`%>%`) does. 


3.	In the Language of Data tutorial, you are tasked with creating a new 
variable in the `email50` dataset, named `num_char_cat`. You were given the
code below:

```{r, eval = FALSE}
# Calculate median number of characters: med_num_char
med_num_char <- median(email50$num_char)

# Create num_char_cat variable in email50
email50_updated <- email50 %>%
  mutate(num_char_cat = if_else(num_char < med_num_char,
                                "[VALUE IF TRUE]",
                                "[VALUE IF FALSE]")
         )
```

Fill in the code with the correct inputs (in place of `[VALUE IF TRUE]` and 
`[VALUE IF FALSE]`) **and** explain what is happening in each line of the code.

4. In the Exploring Categorical Data tutorial, you are tasked with finding the 
proportion of public identity characters that are good. You were given the code
below: 

```{r, eval = FALSE}
comics %>% 
  count(id, align) %>% 
  group_by(id) %>% 
  mutate(prop = n / sum(n)) %>% 
  pivot_wider(id_cols = c(id, align, prop), 
              names_from = align, values_from = prop)
```

Describe what the `group_by()` step does. Why is it necessary to get the 
correct proportions? 

</br>

`r emo::ji("check")` **Check-in 2:** (Due Friday) 

1. What are useful functions for summarizing categorical variables? 

- `count()`
- `mean()`
- `median()`
- `max()`

2. Put the following arguments in the order you would use them as inputs to `ggplot()`.

- `mapping = aes()`
- `data = `
- `geom_bar()`

3. What function do you use if you want to add a new variable to a dataset? 

- `filter()` 
- `mutate()` 
- `count()`
- `select()`

4. What orientation of barchart helps you compare the proportions of one 
categorical variable *across* the levels of another categorical variable? 

- filled (standardized) barchart
- side-by-side (dodged) barcharts
- stacked barcharts
- faceted barcharts

---

# Review from Last Week 

`r emo::ji("microphone")` **Think Out Loud Recording** (Due Sunday)

Record your response to questions exploring a key concept covered during Week 1.
You will be limited to 2 minutes and 30 seconds, to give you an idea of about
how long you will have to describe your thinking in an oral exam. 

Don't think that your recording needs to be perfect, I'd rather you use this as
an opportunity to practice talking through your thinking. Also, I don't want for
this assignment to take up too much of your time, which making lots of
recordings would! 

__Prompt:__ What is dichotomous thinking? How is dichotomous thinking used in
Statistics? What is one benefit and one drawback to dichotomous thinking?


